### 2025-05-17 reportdrafting/test.tex
messing around with some latex, early outline

\documentclass[conference]{IEEEtran}
%Template version as of 6/27/2024

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Neural Networks for Neural Tumors}

\author{
\IEEEauthorblockN{Joe Vogel}
\IEEEauthorblockA{\textit{UC Davis Computer Science} \\
jvogel@ucdavis.edu}

\and
\IEEEauthorblockN{Jonathan Mo}
\IEEEauthorblockA{\textit{UC Davis Computer Science} \\
jtnmo@ucdavis.edu}

\and
\IEEEauthorblockN{Nam Nguyen}
\IEEEauthorblockA{\textit{UC Davis Computer Science} \\
ncnn@ucdavis.edu}

\and
\IEEEauthorblockN{Aryan Karnwal}
\IEEEauthorblockA{\textit{UC Davis Computer Science} \\
akarnwal@ucdavis.edu}

\and
\IEEEauthorblockN{Richard Ho}
\IEEEauthorblockA{\textit{UC Davis Computer Science} \\
rdho@ucdavis.edu}
}

\maketitle

\begin{abstract}
This document is a model and instructions for \LaTeX.
This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
or Math in Paper Title or Abstract.
\end{abstract}

\begin{IEEEkeywords}
component, formatting, style, styling, insert.
\end{IEEEkeywords}

\section{Introduction}
This document is a model and instructions for \LaTeX.
Please observe the conference page limits. For more information about how to become an IEEE Conference author or how to write your paper, please visit   IEEE Conference Author Center website: https://conferences.ieeeauthorcenter.ieee.org/.

\end{document}

### 2025-05-28 reportdrafting/test.tex
implemented things from MQR into final report, needs refactoring

@@
 \usepackage{graphicx}
+\graphicspath{{../images/}}
 \usepackage{textcomp}
@@
     T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+% \setlength{\parindent}{1em}
+% \setlength{\parskip}{0.8em}
 \begin{document}
@@
 \begin{abstract}
-This document is a model and instructions for \LaTeX.
-This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
-or Math in Paper Title or Abstract.
+This report documents the process of making a machine learning model to classify brain tumors using MRI brain scan imagery.
+This project saw the creation of three successful models using the following techniques: CNN, SVM, and Random Forest.
 \end{abstract}
@@
 \begin{IEEEkeywords}
-component, formatting, style, styling, insert.
+Brain tumor classification, Convolutional Neural Network (CNN), MRI brain imagery, Random Forest, Support Vector Machine (SVM), 
 \end{IEEEkeywords}
 
-\section{Introduction}
-This document is a model and instructions for \LaTeX.
-Please observe the conference page limits.
-For more information about how to become an IEEE Conference author or how to write your paper, please visit   IEEE Conference Author Center website: https://conferences.ieeeauthorcenter.ieee.org/.
+\section{Introduction (1-2 pages)}
+Hundreds of thousands of MRI brain scans are performed every day.
+With MRI scans being so intricate, it can sometimes be hard for even a trained professional to determine whether or not a patient has a brain tumor, and even harder to tell which kind.
+
+\subsection{Motivation}
+
+Having a tool that could detect and identify tumors within seconds, and with high accuracy, would save a lot of time and money.
+Not to mention the impact it would have on developing nations who might not have as many specially trained doctors that can readily perform MRI analysis.
+Early detection is absolutely key to diagnosing and treating these tumors.
+
+\subsection{Approach}
+
+A classification machine learning model is the perfect solution to this problem.
+Using a convolutional neural network, we can train a model on MRI images to be able to accurately classify MRI scans into four classifications: glioma tumor, meningioma tumor, pituitary tumor, and no tumor.
+With its complex structure, such a model can detect nuances in the scans that are otherwise unnoticeable to the human eye.
+Given any image of an MRI brain scan, the model should be able to predict these four classes with over 95\% accuracy.
+
+\section{Literature Review (1/2 page)}
+
+\subsection{Brain Tumor Classification in MRI Image Using Convolutional Neural Network}
+
+This paper by Khan et al.
+(2021) begins by discussing brain tumor classification CNN models and their accuracies, which range from 80\% to 98\%.
+The authors point out two key challenges when training such models: first, a lack of properly labeled data, which makes effective training a lot more expensive and imprecise; and second, inconsistencies in image sizes, since a requirement of CNNs is fixed size input.
+Their dataset is divided into three parts: training (for learning model weights), validation (for tuning model parameters), and testing (for final evaluation).
+The authors mentioned preparing the images using an open source computer vision tool called Canny Edge Detection.
+Due to the limited dataset size, data augmentation techniques such as flipping, rotation, and brightness adjustment are applied to increase data variability.
+The model uses the ReLU activation function and stochastic gradient descent for optimization.
+
+Due to limited computing resources, our group will likely rely on a pre-trained model.
+The article mentioned eight layers of neurons so that is a good starting point for us.
+We have put some thought into data pre-processing and will look into Canny Edge Detection as mentioned in the article.
+The article also mentioned the concept of data augmentation, which we will certainly apply.
+Given our small dataset, applying these techniques could help improve generalization on unseen data.
+It was also nice to see familiar terms like ReLU and stochastic gradient descent.
+We will definitely give those a try and see how the techniques affect our model's performance.
+
+\subsection{A Deep CNN Based Multi-Class Classification of Alzheimer's Disease Using MRI}
+
+In this article, Farooq, Anwar, Awais, and Rehman demonstrate their implementation of a 4-way classifier to classify different types of Alzheimer's disease, which is similar to our project since we are classifying our data into 4 sets.
+The authors began with data preprocessing and they utilized a technique known as skull stripping, which is the process of removing non-brain tissues like the skull or fat from MRI images.
+Since our data set consists of MRI images, we can implement skull stripping as well to improve signal-to-noise ratio and reduce input complexity, so it could help handle overfitting.
+Additionally, the article mentions the utilization of data augmentation to increase the data set; they simply flipped the images along the horizontal axis, which is valid due to the left and right symmetry of the brain.
+This method increased their dataset from 355 MRI volumes to 38024 images.
+We can implement the same method in our project because increasing the sample size can combat overfitting, making our model more accurate.
+
+\subsection{Brain Tumor Classification Using Convolutional Neural Networks}
+
+In their 2018 paper, Seetha and Raja demonstrate the effectiveness of Convolutional Neural Networks for automatic brain tumor classification using MRI images.
+Their proposed CNN-based classification approach achieves an impressive 97.5\% accuracy rate while maintaining low computational complexity compared to other methods.
+This could potentially be a performance target we should set on our own model since they used a similar model to us.
+This research is directly relevant to our project as we're also developing a CNN to classify brain MRI images into four categories.
+Their high accuracy rate sets a benchmark we should aspire to, while their focus on computational efficiency aligns with our resource constraints.
+
+The authors implemented a deeper architecture design using small kernels, with neuron weights intentionally kept small to optimize performance.
+Their experimental results demonstrated that their CNN model outperformed other state-of-the-art methods in both accuracy and efficiency.
+Like our approach, they used a dataset containing various brain MRI scans and faced similar challenges with image preprocessing.
+Their optimization of kernel size and weight parameters offers valuable insights for our model development.
+Additionally, our exploration of data augmentation techniques, as mentioned in our mid-quarter report, aligns with common practices in this domain.
+Given that our initial prototype showed signs of overfitting (92.56\% training accuracy but only 42.13\% testing accuracy), we should consider implementing some of their architectural decisions regarding smaller kernels and weight optimization, while continuing with our planned cross-validation and transfer learning approaches to improve generalization.
+
+\section{Dataset Description (1 page)}
+
+The dataset used for this project comes from a group of brilliant students at Savitribai Phule Pune University in India.
+It can be found for free on Kaggle\textsuperscript{1}.
+It is a well respected and reviewed dataset.
+The data consists of 3264 files, which are all images of MRI scans.
+The images are divided into four sets.
+There are three different tumor sets, and one set of “no tumor” scans.
+Considering each image in our dataset is already classified, our model's training is supervised, as opposed to unsupervised, where each image would not have an associated classification to begin with.
+
+\subsection{Preprocessing}
+
+Due to the non-numerical nature of image datasets, our input features come in a different form than they normally would with a numerical dataset.
+Instead of having features such as 'age', 'blood pressure', and so on, we just have an image.
+Technically, the image is a bunch of numbers (pixel intensities), but you get the idea.
+We don't really get to choose which features to train our model on.
+Instead, things like image preprocessing and model hyperparameters are key to training an accurate model.
+
+Preprocessing steps might include things like image resizing, cropping, normalization, and data augmentation (rotating, flipping, zooming, etc.).
+We can even consider techniques like 'skull stripping', used to isolate only the brain, since we do not care about anything else in the scan.
+
+\section{EDA (1 page)}
+
+% image in a column
+% \begin{figure}[!ht]
+%     \centering
+%     \includegraphics[width=2in]{coconut.png}
+%     \caption{this is a coconut}
+%     \label{test image coconut}
+% \end{figure}
+
+% image across two columns
+% \begin{figure*}[!ht]
+%     \centering
+%     \includegraphics[width=3in]{coconut.png}
+%     \caption{this is a coconut}
+%     \label{coconut}
+% \end{figure*}
+
+\begin{figure*}[!ht]
+    \centering
+    \includegraphics[width=4in]{ImagesPerClassTraining.png}
+    \caption{Number of images per class in the training dataset}
+    \label{Number of images per class in the training dataset}
+\end{figure*}
+
+\begin{figure*}[!ht]
+    \centering
+    \includegraphics[width=4in]{ImagesPerClassTesting.png}
+    \caption{Number of images per class in the testing dataset}
+    \label{Number of images per class in the testing dataset}
+\end{figure*}
+
+\begin{figure*}[!ht]
+    \centering
+    \includegraphics[width=4in]{ImageDimensions.png}
+    \caption{Dimensions of the images in the dataset}
+    \label{Dimensions of images in the dataset}
+\end{figure*}
+
+As we can see in Figure 1, there is an even distribution of images among each class (~825), except for the no tumor class (395).
+This is because the images without tumors have much less variance.
+It should be much easier for the model to detect tumor vs no tumor, so we do not expect the lower number of images in the no tumor class to have a negative effect on the model's ability to classify images as such.
+However, if we do see poor results in that classification specifically, we can try changing all classes to train on 395 images, to see if that helps with the class-specific overfitting.
+
+In our test dataset seen in Figure 2, we have a much more uneven distribution among each class.
+This of course won't affect our model's ability, since it is not being trained on this data, but some results could be skewed considering the relatively low number of pituitary images.
+It could be useful to test against more pituitary imagery to ensure the model can handle all variations of scans.
+We could perform data augmentation and create more data by rotating, zooming, and flipping our existing data.
+Testing against more data could give us more confidence in our model, if scores remain high.
+
+We can see in Figure 3 that our dataset is comprised of images of varying sizes.
+There seems to be a concentration around the 200x200 and 500x500 sizes.
+With this information, we know we will have to perform some image pre-processing to convert all images into the same size so the model has consistent input.
+A common image size is 256x256.
+
+\section{Methodology (1-2 pages)}
+
+\subsection{Feature Selection}
+As mentioned in our dataset description, we don't have labels for our input features like a numerical dataset would have.
+We only have our images, which are technically a bunch of pixel intensity values, but it means we can't select certain features to train our dataset on.
+Instead, we can perform some image preprocessing and hyperparameter tuning to increase our model's accuracy.
+
+If we look at Figure 3 again, we see all of our images have different dimensions.
+As a preprocessing step, we will resize them all to be 256x256 pixels, so that our model has consistent input.
+We then normalize each pixel's value from its 3-channel RGB form.
+The resizing step is not as simple as it sounds though, we must be careful because shrinking or enlarging our images can create noise in our data.
+Based on future results, we might try a few different image sizes to see which is best.
+
+Our training dataset consists of 2870 MRI scans.
+If we want to give our model more images to train on, we can perform data augmentation.
+We can rotate, flip, and zoom our images to create 'artificial' data that will hopefully prevent overfitting and allow for our model to capture underlying patterns better.
+We also hope this will allow for our model to predict against a larger variety of testing data (i.e.
+rotated images).
+
+The hyperparameters of our model are crucial to achieving a good fit. In our first iteration of the model, some of the hyperparameters we chose were as follows:
+%TODO --- fix table
+- Image size: 256x256 pixels		- Batch size: 32		- Epochs: 30	
+- Loss:  Categorical Cross Entropy 	- Metrics: Accuracy		- Dropout Rate: 0.25 and 0.50
+We plan on fine tuning these parameters with techniques like grid search as we continue to optimize our model.
+
+\subsection{Early Model Development}
+
+We have already built an early prototype of a CNN model with 4 convolutional blocks, followed by fully connected layers using Keras Sequential API.
+We trained our model over 30 epochs, with each epoch giving us higher accuracy, peaking at 92.56\% after the last epoch.
+However, when tested against our testing data, the model performs quite poorly.
+We see an average accuracy of 42.13\%, indicating that the model is overfitting quite heavily on the training data.
+
+To improve upon our first model, we'll apply k-fold cross validation and tune hyperparameters such as dropout rate, number of layers/neurons, etc.
+We could start with some manual tuning, and eventually try implementing grid search to tune our hyperparameters.
+We also plan to use transfer learning to reduce the number of learned parameters and implicitly filter irrelevant features, which can greatly reduce overfitting risks.
+Going forward, we'll use PyTorch, which gives more flexibility in controlling forward pass, backpropagation, and optimization, to develop a better CNN model.
+
+\section{Experimental Results (2 pages)}
+
+\section{Evaluation (1-2 pages)}
+
+\section{Conclusion (1/2 page)}
+
+\section{Our Team (1 page)}
+The link to our GitHub repository for this project can be foun at: https://github.com/Coco501/Neural-Networks-for-Neural-Tumors
+
+project roadmap? 
+%TODO --- roadmap?
+
+\subsection{Roles and Contributions}
+
+Joe: As the team leader, Joe has taken charge of organizing meetings, assigning tasks, and the general oversight over the project.
+In the coming weeks, Joe will be focusing on Exploratory Data Analysis, Literature Review, and Model Optimization.
+
+Nam: Nam took charge of making the first prototype of our model.
+Having this model as a starting and reference point is absolutely crucial in the development of subsequent models.
+In the next few weeks, Nam will be researching different frameworks and libraries to develop other CNN models to see which best fit our dataset.
+
+Jonathan: Jonathan took charge of literature review and researched techniques that can be implemented to the model.
+In the next few weeks, Jonathan will be focusing on writing the proposed methodology for the final report.
+
+Richard: Richard has completed an article for literature review.
+Over the coming weeks, he will work towards exploratory data analysis (EDA) and preparing the dataset for analysis through data pre-processing.
+
+Aryan: Aryan completed a literature review and analyzed research papers to identify techniques we can use to address the models overfitting.
+In the next few weeks Aryan will continue working to improve the models accuracy while also working on documentation and reports.
+
+\begin{thebibliography}{00}
+
+\bibitem{Khan2021}
+H. A. Khan, W. Jue, M. Mushtaq, and M. U. Mushtaq, “Brain tumor classification in MRI image using convolutional neural network,” *Math. Biosci. Eng.*, 2021.
+
+\bibitem{Kumar2014}
+R. S. Kumar and M. Karnan, “Review of MRI image classification techniques,” *Int. J. Res. Stud. Comput. Sci. Eng.*, vol. 1, no. 1, pp. 21-28, 2014.
+
+\bibitem{Farooq2017}
+A. Farooq, S. Anwar, M. Awais, and S. Rehman, “A deep CNN based multi-class classification of Alzheimer's disease using MRI,” in *Proc. IEEE Int. Conf. Imaging Syst. Tech. (IST)*, 2017, pp. 1-6.
+
+\bibitem{Hasan2019}
+A. M. Hasan, H. A. Jalab, F. Meziane, H. Kahtan, and A. S. Al-Ahmad, “Combining deep and handcrafted image features for MRI brain scan classification,” *IEEE Access*, vol. 7, pp. 79959-79967, 2019.
+
+\bibitem{Meier2020}
+R.
+Meier, A.
+P.
+de Mortanges, R.
+Wiest, and U.
+Knecht, “Exploratory analysis of qualitative MR imaging features for the differentiation of glioblastoma and brain metastases,” *Front.
+Oncol.*, vol.
+10, p.
+581037, 2020.
+
+\bibitem{Seetha2018}
+J. Seetha and S. S. Raja, “Brain tumor classification using convolutional neural networks,” *Biomed. Pharmacol. J.*, vol. 11, no. 3, p. 1457, 2018.
+
+\end{thebibliography}
 

### 2025-06-01 latex/draft.tex
renamed
minor adjustments so far

--- 2025-05-28 reportdrafting/test.tex
@@
 \usepackage{graphicx}
-\graphicspath{{../images/}}
+\graphicspath{{images/}}
 \usepackage{textcomp}
@@
 \author{
-\IEEEauthorblockN{Joe Vogel}
+\IEEEauthorblockN{\large Joe Vogel}
 \IEEEauthorblockA{\textit{UC Davis Computer Science} \\
@@
 \and
-\IEEEauthorblockN{Jonathan Mo}
+\IEEEauthorblockN{\large Nam Nguyen}
+\IEEEauthorblockA{\textit{UC Davis Computer Science} \\
+ncnn@ucdavis.edu}
+
+\and
+\IEEEauthorblockN{\large Jonathan Mo}
 \IEEEauthorblockA{\textit{UC Davis Computer Science} \\
@@
 \and
-\IEEEauthorblockN{Nam Nguyen}
-\IEEEauthorblockA{\textit{UC Davis Computer Science} \\
-ncnn@ucdavis.edu}
-
-\and
-\IEEEauthorblockN{Aryan Karnwal}
+\IEEEauthorblockN{\large Aryan Karnwal}
 \IEEEauthorblockA{\textit{UC Davis Computer Science} \\
@@
 \and
-\IEEEauthorblockN{Richard Ho}
+\IEEEauthorblockN{\large Richard Ho}
 \IEEEauthorblockA{\textit{UC Davis Computer Science} \\
@@
 \begin{abstract}
-This report documents the process of making a machine learning model to classify brain tumors using MRI brain scan imagery.
-This project saw the creation of three successful models using the following techniques: CNN, SVM, and Random Forest.
+This report documents the process of creating a machine learning model that classifies brain tumors using MRI brain scan imagery.
+This project saw the creation of three successful models using the following algorithms: Convolutional Neural Network, Support Vector Machine, and Random Forest.
 \end{abstract}
@@
 \begin{IEEEkeywords}
-Brain tumor classification, Convolutional Neural Network (CNN), MRI brain imagery, Random Forest, Support Vector Machine (SVM), 
+Brain tumor classification, Convolutional Neural Network (CNN), MRI brain imagery, Random Forest (RF), Support Vector Machine (SVM). 
 \end{IEEEkeywords}
 
-\section{Introduction (1-2 pages)}
+\large 
+\section{\large Introduction (1-2 pages)}
 Hundreds of thousands of MRI brain scans are performed every day.
@@
 
-\subsection{Motivation}
-
+\subsection{\large Motivation}
 Having a tool that could detect and identify tumors within seconds, and with high accuracy, would save a lot of time and money.
@@
 
-\subsection{Approach}
+\subsection{\large Approach}
+A classification machine learning model is the perfect solution to this problem.
+Using a convolutional neural network and an image dataset, we can train a model to accuractely classify brain tumors.
+With its complex neural structure, such a model can detect nuances in the scans that are otherwise unnoticeable to the human eye.
+This project used a dataset that focused on three brain tumor types: glioma, meningioma, pituitary.
+The model will also be able to predict scans that have "no tumor".
+Given an image of an MRI brain scan containing one of the four classifications, the model should predict the tumor class with over 95\% accuracy.
 
-A classification machine learning model is the perfect solution to this problem.
-Using a convolutional neural network, we can train a model on MRI images to be able to accurately classify MRI scans into four classifications: glioma tumor, meningioma tumor, pituitary tumor, and no tumor.
-With its complex structure, such a model can detect nuances in the scans that are otherwise unnoticeable to the human eye.
-Given any image of an MRI brain scan, the model should be able to predict these four classes with over 95\% accuracy.
+\section{\large Literature Review (1/2 page)}
 
-\section{Literature Review (1/2 page)}
-
-\subsection{Brain Tumor Classification in MRI Image Using Convolutional Neural Network}
+\subsection{\large Brain Tumor Classification in MRI Image Using Convolutional Neural Network$^{1}$}
 
@@
 
-\subsection{A Deep CNN Based Multi-Class Classification of Alzheimer's Disease Using MRI}
+\subsection{\large A Deep CNN Based Multi-Class Classification of Alzheimer's Disease Using MRI$^{2}$}
 
@@
 
-\subsection{Brain Tumor Classification Using Convolutional Neural Networks}
+\subsection{\large Brain Tumor Classification Using Convolutional Neural Networks$^{3}$}
 
@@
 
-\section{Dataset Description (1 page)}
+\section{\large Dataset Description (1 page)}
 
 The dataset used for this project comes from a group of brilliant students at Savitribai Phule Pune University in India.
-It can be found for free on Kaggle\textsuperscript{1}.
+It can be found for free on Kaggle\textsuperscript{4}.
 It is a well respected and reviewed dataset.
@@
 
-\subsection{Preprocessing}
+\subsection{\large Preprocessing}
 
@@
 
-\section{EDA (1 page)}
+\section{\large EDA (1 page)}
 
@@
     \includegraphics[width=4in]{ImagesPerClassTraining.png}
-    \caption{Number of images per class in the training dataset}
+    \caption{\large Number of images per class in the training dataset}
     \label{Number of images per class in the training dataset}
@@
     \includegraphics[width=4in]{ImagesPerClassTesting.png}
-    \caption{Number of images per class in the testing dataset}
+    \caption{\large Number of images per class in the testing dataset}
     \label{Number of images per class in the testing dataset}
@@
     \includegraphics[width=4in]{ImageDimensions.png}
-    \caption{Dimensions of the images in the dataset}
+    \caption{\large Dimensions of the images in the dataset}
     \label{Dimensions of images in the dataset}
@@
 
-\section{Methodology (1-2 pages)}
+\section{\large Methodology (1-2 pages)}
 
@@
 
-\subsection{Early Model Development}
+\subsection{\large Early Model Development}
 
@@
 
-\section{Experimental Results (2 pages)}
+\section{\large Experimental Results (2 pages)}
 
-\section{Evaluation (1-2 pages)}
+\section{\large Evaluation (1-2 pages)}
 
-\section{Conclusion (1/2 page)}
+\section{\large Conclusion (1/2 page)}
 
-\section{Our Team (1 page)}
+\section{\large Our Team (1 page)}
 The link to our GitHub repository for this project can be foun at: https://github.com/Coco501/Neural-Networks-for-Neural-Tumors
@@
 
-\subsection{Roles and Contributions}
+\subsection{\large Roles and Contributions}
 
@@
 
+\bibitem{Farooq2017}
+A. Farooq, S. Anwar, M. Awais, and S. Rehman, “A deep CNN based multi-class classification of Alzheimer's disease using MRI,” in *Proc. IEEE Int. Conf. Imaging Syst. Tech. (IST)*, 2017, pp. 1-6.
+
+\bibitem{Seetha2018}
+J. Seetha and S. S. Raja, “Brain tumor classification using convolutional neural networks,” *Biomed. Pharmacol. J.*, vol. 11, no. 3, p. 1457, 2018.
+
+\bibitem{Sartaj2020}
+Sartaj. “Brain Tumor Classification (MRI).” Kaggle, 24 May 2020, www.kaggle.com/datasets/sartajbhuvaji/brain-tumor-classification-mri.
+
 \bibitem{Kumar2014}
 R. S. Kumar and M. Karnan, “Review of MRI image classification techniques,” *Int. J. Res. Stud. Comput. Sci. Eng.*, vol. 1, no. 1, pp. 21-28, 2014.
-
-\bibitem{Farooq2017}
-A. Farooq, S. Anwar, M. Awais, and S. Rehman, “A deep CNN based multi-class classification of Alzheimer's disease using MRI,” in *Proc. IEEE Int. Conf. Imaging Syst. Tech. (IST)*, 2017, pp. 1-6.
 
@@
 
-\bibitem{Seetha2018}
-J. Seetha and S. S. Raja, “Brain tumor classification using convolutional neural networks,” *Biomed. Pharmacol. J.*, vol. 11, no. 3, p. 1457, 2018.
-
 \end{thebibliography}

### 2025-06-01 latex/draft.tex
minor adjustments

@@
 
+% old author block
+% \author{
+% \IEEEauthorblockN{\large Joe Vogel}
+% \IEEEauthorblockA{\textit{UC Davis Computer Science} \\
+% jvogel@ucdavis.edu}
+
+% \and
+% \IEEEauthorblockN{\large Nam Nguyen}
+% \IEEEauthorblockA{\textit{UC Davis Computer Science} \\
+% ncnn@ucdavis.edu}
+
+% \and
+% \IEEEauthorblockN{\large Jonathan Mo}
+% \IEEEauthorblockA{\textit{UC Davis Computer Science} \\
+% jtnmo@ucdavis.edu}
+
+% \and
+% \IEEEauthorblockN{\large Aryan Karnwal}
+% \IEEEauthorblockA{\textit{UC Davis Computer Science} \\
+% akarnwal@ucdavis.edu}
+
+% \and
+% \IEEEauthorblockN{\large Richard Ho}
+% \IEEEauthorblockA{\textit{UC Davis Computer Science} \\
+% rdho@ucdavis.edu}
+% }
+
 \author{
-\IEEEauthorblockN{\large Joe Vogel}
-\IEEEauthorblockA{\textit{UC Davis Computer Science} \\
-jvogel@ucdavis.edu}
-
-\and
-\IEEEauthorblockN{\large Nam Nguyen}
-\IEEEauthorblockA{\textit{UC Davis Computer Science} \\
-ncnn@ucdavis.edu}
-
-\and
-\IEEEauthorblockN{\large Jonathan Mo}
-\IEEEauthorblockA{\textit{UC Davis Computer Science} \\
-jtnmo@ucdavis.edu}
-
-\and
-\IEEEauthorblockN{\large Aryan Karnwal}
-\IEEEauthorblockA{\textit{UC Davis Computer Science} \\
-akarnwal@ucdavis.edu}
-
-\and
-\IEEEauthorblockN{\large Richard Ho}
-\IEEEauthorblockA{\textit{UC Davis Computer Science} \\
-rdho@ucdavis.edu}
+\centering
+\begin{minipage}[t]{0.9\textwidth}
+  \centering
+  \begin{tabular}{ccc}
+    Joe Vogel & Nam Nguyen & Jonathan Mo \\
+    \textit{UC Davis Computer Science} & \textit{UC Davis Computer Science} & \textit{UC Davis Computer Science} \\
+    jvogel@ucdavis.edu & ncnn@ucdavis.edu & jtnmo@ucdavis.edu \\
+    \vspace{0.6em}
+  \end{tabular}
+  \begin{tabular}{cc}
+    Aryan Karnwal & Richard Ho \\
+    \textit{UC Davis Computer Science} & \textit{UC Davis Computer Science} \\
+    akarnwal@ucdavis.edu & rdho@ucdavis.edu \\
+  \end{tabular}
+\end{minipage}
 }
@@
 The images are divided into four sets.
-There are three different tumor sets, and one set of “no tumor” scans.
+There are three different tumor types, and one set of “no tumor” scans.
 Considering each image in our dataset is already classified, our model's training is supervised, as opposed to unsupervised, where each image would not have an associated classification to begin with.

### 2025-06-01 latex/draft.tex
more report drafting

@@
 
-% image in a column
+% EXAMPLE image in a column
 % \begin{figure}[!ht]
@@
 
-% image across two columns
+% EXAMPLE image across two columns
 % \begin{figure*}[!ht]
@@
 
+% These two won't stay in the same row
+% \begin{figure*}[!ht]
+%     \centering
+%     \includegraphics[width=1in]{ImagesPerClassTraining.png}
+%     \caption{\large Images in the training dataset}
+%     \label{Number of images per class in the training dataset}
+% \end{figure*}
+
+% \begin{figure*}[!ht]
+%     \centering
+%     \includegraphics[width=1in]{ImagesPerClassTesting.png}
+%     \caption{\large Images in the testing dataset}
+%     \label{Number of images per class in the testing dataset}
+% \end{figure*}
+
+% Both figures in one row
 \begin{figure*}[!ht]
     \centering
-    \includegraphics[width=4in]{ImagesPerClassTraining.png}
-    \caption{\large Number of images per class in the training dataset}
-    \label{Number of images per class in the training dataset}
+    \includegraphics[width=0.49\textwidth]{ImagesPerClassTraining.png}
+    \hfill
+    \includegraphics[width=0.49\textwidth]{ImagesPerClassTesting.png}
+    \caption{\large Number of images per class in the training (left) and testing (right) datasets}
+    \label{fig:images-per-class}
 \end{figure*}
 
-\begin{figure*}[!ht]
-    \centering
-    \includegraphics[width=4in]{ImagesPerClassTesting.png}
-    \caption{\large Number of images per class in the testing dataset}
-    \label{Number of images per class in the testing dataset}
-\end{figure*}
+As we can see in Figure 1 (left), there is an even distribution of images among each class (~825), except for the no tumor class (395).
+This is because the images without tumors have much less variance.
+It should be much easier for the model to detect tumor vs no tumor, so we do not expect the lower number of images in the no tumor class to have a negative effect on the model's ability to classify images as such.
+However, if we do see poor results in that classification specifically, we can try changing all classes to train on 395 images, to see if that helps with the class-specific overfitting.
+
+In our test dataset seen in Figure 1 (right), we have a much more uneven distribution among each class.
+This of course won't affect our model's ability, since it is not being trained on this data, but some results could be skewed considering the relatively low number of pituitary images.
+It could be useful to test against more pituitary imagery to ensure the model can handle all variations of scans.
+We could perform data augmentation and create more data by rotating, zooming, and flipping our existing data.
+Testing against more data could give us more confidence in our model, if scores remain high.
 
@@
 
-As we can see in Figure 1, there is an even distribution of images among each class (~825), except for the no tumor class (395).
-This is because the images without tumors have much less variance.
-It should be much easier for the model to detect tumor vs no tumor, so we do not expect the lower number of images in the no tumor class to have a negative effect on the model's ability to classify images as such.
-However, if we do see poor results in that classification specifically, we can try changing all classes to train on 395 images, to see if that helps with the class-specific overfitting.
-
-In our test dataset seen in Figure 2, we have a much more uneven distribution among each class.
-This of course won't affect our model's ability, since it is not being trained on this data, but some results could be skewed considering the relatively low number of pituitary images.
-It could be useful to test against more pituitary imagery to ensure the model can handle all variations of scans.
-We could perform data augmentation and create more data by rotating, zooming, and flipping our existing data.
-Testing against more data could give us more confidence in our model, if scores remain high.
-
-We can see in Figure 3 that our dataset is comprised of images of varying sizes.
+We can see in Figure 2 that our dataset is comprised of images of varying sizes.
 There seems to be a concentration around the 200x200 and 500x500 sizes.
@@
 
-\subsection{Feature Selection}
+\subsection{\large Feature Selection}
 As mentioned in our dataset description, we don't have labels for our input features like a numerical dataset would have.
@@
 \section{\large Experimental Results (2 pages)}
+Talk about trying three different models, different pretrained models like ResNet, (the other one you tried first Nam I forgot what it was)
+
+Talk about fine-tuning, grid search for hyperparameter fine-tuning, etc.
+
+Nam, Elaborate on this: "Alright guys, I finally figured out the problem.
+I've been using different pre-trained models but keep ending up with a lot of false negatives for the class glioma.
+I reviewed another notebook using the same dataset.
+Comparing it with our reference notebook, I notice that they both initially combined the training and testing folders, shuffled and split again.
+On the other hand, I'd been using the dataset as it is split on Kaggle.
+I believe there must be significant patterns in the testing folders on Kaggle that are not present in the training ones, especially the glioma-tumor class.
+So, merging them together and splitting again is the optimal solution here.
+Btw, I also decided to use Keras Function API instead of Sequential API (still same network architecture) and it turned out to be slightly better.
+I run the model several times and the accuracy is averagely around 92\% to 95\%"
 
 \section{\large Evaluation (1-2 pages)}
+Include all 3 models and their confusion matrix results, do not include the actual confusion matrix, only the f1 scores, precision, etc.
 
 \section{\large Conclusion (1/2 page)}
+Talk about how we chose CNN as our final model and reiterate how it performs
 
 \section{\large Our Team (1 page)}
-The link to our GitHub repository for this project can be foun at: https://github.com/Coco501/Neural-Networks-for-Neural-Tumors
+The link to our GitHub\textsuperscript{8} repository for this project can be found in the references of this report.
 
-project roadmap? 
+project roadmap? Reference the first one pager outline assignment we did on google doc
 %TODO --- roadmap?
+\subsection{\large Demonstration}
+include youtube video of our demo
 
@@
 
+\bibitem{Vogel2025}
+J. Vogel, N. Nguyen, J. Mo, A. Karnwal, R. Ho, "Neural Networks for Neural Tumors". GitHub. https://github.com/Coco501/Neural-Networks-
+for-Neural-Tumors
+
 \end{thebibliography}

### 2025-06-01 src/frontend/app.py
frontend files added (keras file not included)

@@ -1,0 +1,228 @@
+import gradio as gr
+import tensorflow as tf
+import numpy as np
+from PIL import Image
+import cv2
+import os
+
+os.environ['TF_CPP_MIN_LOG_LEVEL'] = '2'
+
+def load_model_safe():
+    model_paths = [
+        'CNN_model_compatible.keras',
+        'CNN_model.keras',
+        'brain_tumor_model',
+    ]
+    
+    for model_path in model_paths:
+        if os.path.exists(model_path):
+            try:
+                print(f"Attempting to load model from: {model_path}")
+                
+                if model_path.endswith('.keras'):
+                    model = tf.keras.models.load_model(model_path, compile=False)
+                    model.compile(
+                        optimizer='adam',
+                        loss='categorical_crossentropy',
+                        metrics=['accuracy']
+                    )
+                    print("Model recompiled successfully")
+                else:
+                    model = tf.keras.models.load_model(model_path)
+                
+                print(f"Model loaded successfully from: {model_path}")
+                return model
+                
+            except Exception as e:
+                print(f"Failed to load from {model_path}: {e}")
+                continue
+    
+    try:
+        print("Attempting manual model reconstruction...")
+        if os.path.exists('model_architecture.json') and os.path.exists('model_weights.h5'):
+            from tensorflow.keras.models import model_from_json
+            
+            with open('model_architecture.json', 'r') as json_file:
+                model_json = json_file.read()
+            
+            model = model_from_json(model_json)
+            model.load_weights('model_weights.h5')
+            model.compile(
+                optimizer='adam',
+                loss='categorical_crossentropy',
+                metrics=['accuracy']
+            )
+            
+            print("Model reconstructed and compiled successfully!")
+            return model
+    except Exception as e:
+        print(f"Manual reconstruction failed: {e}")
+    
+    print("All model loading attempts failed!")
+    return None
+
+model = load_model_safe()
+
+if model is not None:
+    print(f"Model input shape: {model.input_shape}")
+    print(f"Model output shape: {model.output_shape}")
+    print(f"Model summary:")
+    model.summary()
+else:
+    print("Model failed to load!")
+
+class_labels = ['glioma', 'meningioma', 'no_tumor', 'pituitary']
+
+def preprocess_image(image):
+    from tensorflow.keras.applications.imagenet_utils import preprocess_input
+    
+    if isinstance(image, Image.Image):
+        img_array = np.array(image)
+    else:
+        img_array = image
+    
[…]

### 2025-06-01 latex/draft.tex
evaluation and conclusion draft

@@
 
-This paper by Khan et al.
-(2021) begins by discussing brain tumor classification CNN models and their accuracies, which range from 80\% to 98\%.
-The authors point out two key challenges when training such models: first, a lack of properly labeled data, which makes effective training a lot more expensive and imprecise; and second, inconsistencies in image sizes, since a requirement of CNNs is fixed size input.
-Their dataset is divided into three parts: training (for learning model weights), validation (for tuning model parameters), and testing (for final evaluation).
-The authors mentioned preparing the images using an open source computer vision tool called Canny Edge Detection.
-Due to the limited dataset size, data augmentation techniques such as flipping, rotation, and brightness adjustment are applied to increase data variability.
-The model uses the ReLU activation function and stochastic gradient descent for optimization.
+Khan et al.
+(2021) examine convolutional neural network (CNN) models for brain tumor classification, reporting classification accuracies ranging from 80\% to 98\%.
+The authors highlight two primary challenges in training such models: (1) the scarcity of accurately labeled data, which increases training complexity and cost, and (2) inconsistency in image dimensions, given that CNNs require fixed-size input.
+Their dataset is partitioned into training, validation, and testing sets, serving model training, hyperparameter tuning, and final evaluation, respectively.
+Image pre-processing is performed using Canny Edge Detection, an open-source computer vision technique.
+To address the limited dataset size, the authors employ data augmentation methods including flipping, rotation, and brightness adjustments to enhance variability.
+The model utilizes the ReLU activation function and stochastic gradient descent (SGD) for optimization.
 
-Due to limited computing resources, our group will likely rely on a pre-trained model.
-The article mentioned eight layers of neurons so that is a good starting point for us.
-We have put some thought into data pre-processing and will look into Canny Edge Detection as mentioned in the article.
-The article also mentioned the concept of data augmentation, which we will certainly apply.
-Given our small dataset, applying these techniques could help improve generalization on unseen data.
-It was also nice to see familiar terms like ReLU and stochastic gradient descent.
-We will definitely give those a try and see how the techniques affect our model's performance.
+Given our limited computational resources, we plan to utilize a pre-trained model.
+The eight-layer architecture discussed in the paper provides a useful baseline.
+We are also considering Canny Edge Detection for pre-processing, as suggested by the authors.
+Furthermore, we intend to implement data augmentation techniques to improve generalization on our small dataset.
+The use of ReLU and SGD aligns well with our current understanding, and we aim to assess their impact on our model’s performance.
 
@@
 \section{\large Evaluation (1-2 pages)}
-Include all 3 models and their confusion matrix results, do not include the actual confusion matrix, only the f1 scores, precision, etc.
+In this project, we used three learning models (CNN, SVM, and RF) and compared the results of each. The dataset comprises four categories: glioma, meningioma, pituitary, and no tumor.
+
+The CNN model demonstrated the highest overall accuracy at 96.68\%, followed closely by SVM at 95.66\%, and RF at 94.90\%.
+Macro and weighted F1-scores mirrored this trend, with CNN achieving 0.97, SVM 0.96, and RF 0.95 in both averages.
+These scores indicate that CNN not only achieves better overall classification performance but also maintains consistency across classes in terms of precision and recall.
+
+Diving into per-class specifics, we observe that for the label "glioma", CNN exhibited the highest precision (0.98) and recall (0.93), indicating strong identification capability with few false positives and a relatively low false negative rate.
+SVM matched CNN in precision (0.98) but had a lower recall (0.91), while RF also had a precision (0.98) but the lowest score in recall at 0.89.
+This suggests that both SVM and RF are more likely to miss "glioma" cases compared to CNN while matching its precision.
+
+In the case of "meningioma", while CNN had the best precision (0.93), SVM and RF had equal or superior recall (0.97 and 0.98 respectively) compared to CNN (0.97).
+SVM and RF had reduced precision (0.89 and 0.87) which is a notable difference from CNN's aforementioned score.
+This pattern indicates that SVM and RF are more sensitive in detecting "meningioma" but may yield more false positives than CNN.
+
+In classifying "no tumor", RF achieved perfect precision (1.00), meaning that every time RF predicted "no tumor", the model was correct.
+However, its recall (0.93) was the lowest among the three, pointing to a higher rate of false negatives.
+CNN balanced both precision (0.97) and recall (0.98) well, outperforming both alternatives in overall reliability for this class.
+
+For the label "pituitary", all three models performed strongly, with CNN leading in both precision and recall (0.99 each).
+SVM and RF followed closely with both metrics at 0.98 and 0.97-0.98, respectively, indicating that all models are highly effective in classifying "pituitary" tumors.
 
 \section{\large Conclusion (1/2 page)}
-Talk about how we chose CNN as our final model and reiterate how it performs
+The evaluation highlight CNN's strength in handling complex, high-dimensional image data, likely due to its hierarchical feature extraction capability.
+Its superior balance in both precision and recall across all tumor types supports its suitability for real-world clinical deployment where both false positives and false negatives carry significant risk.
+
+SVM performed comparably well, especially in the detection of meningioma and pituitary tumors, and may offer a lighter-weight alternative in environments with limited computational resources.
+However, it showed reduced precision in some classes, making it slightly less robust overall.
+
+RF, while achieving remarkable precision, particularly for the "no tumor" class, exhibited reduced recall across multiple classes.
+This suggests a tendency to make fewer false positives but at the expense of missing actual tumor cases, which is a very critical and life-threatening issue in medical diagnostics.
 

### 2025-06-02 latex/draft.tex
more report drafting

@@
 \large 
-\section{\large Introduction (1-2 pages)}
-Hundreds of thousands of MRI brain scans are performed every day.
-With MRI scans being so intricate, it can sometimes be hard for even a trained professional to determine whether or not a patient has a brain tumor, and even harder to tell which kind.
+\section{\large Introduction (Front page)}
+As a team of five young computer science students with no prior machine learning experience, we ambitously decided to tackle the real world challenge of tumor classification for our \textit{ECS 171 - Machine Learning} final project at UC Davis.
+Through hours and hours of research and trial and error, we learned a lot, and we have a lot to share.
+
+We hope this report serves not only as an informative summary of our findings, but also as motivation for your own projects.
+We faced many challenges during this project, but in the end created something we are all very proud of, despite our lack of experience.
 
 \subsection{\large Motivation}
+
+Hundreds of thousands of MRI brain scans are performed every day.
+With the scans being so intricate, it can sometimes be hard for even a trained professional to determine whether or not a patient has a brain tumor, and even harder to tell which kind.
+
 Having a tool that could detect and identify tumors within seconds, and with high accuracy, would save a lot of time and money.
 Not to mention the impact it would have on developing nations who might not have as many specially trained doctors that can readily perform MRI analysis.
-Early detection is absolutely key to diagnosing and treating these tumors.
+With lives on the line, early detection is absolutely key to diagnosing and treating these tumors.
 
 \subsection{\large Approach}
-A classification machine learning model is the perfect solution to this problem.
+A classification machine learning model is a great approach to this problem.
 Using a convolutional neural network and an image dataset, we can train a model to accuractely classify brain tumors.
-With its complex neural structure, such a model can detect nuances in the scans that are otherwise unnoticeable to the human eye.
+With its complex neural structure, such a model can detect nuances and patterns in the scans that are otherwise hard to spot.
 This project used a dataset that focused on three brain tumor types: glioma, meningioma, pituitary.
-The model will also be able to predict scans that have "no tumor".
-Given an image of an MRI brain scan containing one of the four classifications, the model should predict the tumor class with over 95\% accuracy.
+Along with the three tumor types, the model should also be able to classify scans as "no tumor".
+The aim for this model was to predict tumor classes with over 95\% accuracy.
 
-\section{\large Literature Review (1/2 page)}
+\section{\large Literature Review (Complete)}
 
@@
 Khan et al.
-(2021) examine convolutional neural network (CNN) models for brain tumor classification, reporting classification accuracies ranging from 80\% to 98\%.
+(2021) examine convolutional neural network models for brain tumor classification, reporting classification accuracies ranging from 80\% to 98\%.
 The authors highlight two primary challenges in training such models: (1) the scarcity of accurately labeled data, which increases training complexity and cost, and (2) inconsistency in image dimensions, given that CNNs require fixed-size input.
@@
 Image pre-processing is performed using Canny Edge Detection, an open-source computer vision technique.
-To address the limited dataset size, the authors employ data augmentation methods including flipping, rotation, and brightness adjustments to enhance variability.
-The model utilizes the ReLU activation function and stochastic gradient descent (SGD) for optimization.
+To address the limited dataset size, the authors employ data augmentation methods such as flipping, rotation, and brightness adjustments to enhance variability.
+The model utilizes the ReLU activation function and stochastic gradient descent for optimization.
 
@@
 The eight-layer architecture discussed in the paper provides a useful baseline.
-We are also considering Canny Edge Detection for pre-processing, as suggested by the authors.
+The pre-processing techniques mentioned are also great starting points for us, and let us know what we should look out for.
 Furthermore, we intend to implement data augmentation techniques to improve generalization on our small dataset.
-The use of ReLU and SGD aligns well with our current understanding, and we aim to assess their impact on our model’s performance.
+The use of ReLU and stochastic gradient descent aligns well with our current understanding, and we aim to assess their impact on our model's performance.
 
@@
 
-In their 2018 paper, Seetha and Raja demonstrate the effectiveness of Convolutional Neural Networks for automatic brain tumor classification using MRI images.
+In this publication, authors Seetha and Raja demonstrate the effectiveness of convolutional neural networks for brain tumor classification.
 Their proposed CNN-based classification approach achieves an impressive 97.5\% accuracy rate while maintaining low computational complexity compared to other methods.
-This could potentially be a performance target we should set on our own model since they used a similar model to us.
-This research is directly relevant to our project as we're also developing a CNN to classify brain MRI images into four categories.
-Their high accuracy rate sets a benchmark we should aspire to, while their focus on computational efficiency aligns with our resource constraints.
+Seeing such a high accuracy be achieved is both inspiring and motivating for us.
+Considering their model also predicts four tumor types, their high accuracy rate sets a benchmark we should aspire to, and their focus on computational efficiency aligns with our resource constraints.
 
@@
 Their optimization of kernel size and weight parameters offers valuable insights for our model development.
-Additionally, our exploration of data augmentation techniques, as mentioned in our mid-quarter report, aligns with common practices in this domain.
-Given that our initial prototype showed signs of overfitting (92.56\% training accuracy but only 42.13\% testing accuracy), we should consider implementing some of their architectural decisions regarding smaller kernels and weight optimization, while continuing with our planned cross-validation and transfer learning approaches to improve generalization.
+Given that our initial prototype showed signs of overfitting (93\% training accuracy with 42\% testing accuracy), we should consider implementing some of the suggested architectural decisions regarding smaller kernels and weight optimizations, while continuing with our planned transfer learning approach to improve generalization.
 
@@
 It is a well respected and reviewed dataset.
-The data consists of 3264 files, which are all images of MRI scans.
-The images are divided into four sets.
+The dataset consists of 3264 files, which are all images of MRI scans.
+The images are divided into four categories.
 There are three different tumor types, and one set of “no tumor” scans.
-Considering each image in our dataset is already classified, our model's training is supervised, as opposed to unsupervised, where each image would not have an associated classification to begin with.
+Because each image in our dataset is already classified, our model's training is considered supervised, as opposed to unsupervised, where each training image would not have an associated classification.
 
@@
 Due to the non-numerical nature of image datasets, our input features come in a different form than they normally would with a numerical dataset.
-Instead of having features such as 'age', 'blood pressure', and so on, we just have an image.
-Technically, the image is a bunch of numbers (pixel intensities), but you get the idea.
-We don't really get to choose which features to train our model on.
+Instead of having numerically represented features such as 'age', 'blood pressure', and so on, our dataset only has images.
+Technically, each image is a set of numbers (pixel intensities), but the idea remains.
+We don't really have a standard feature selection process, or get to choose which features to train our model on.
 Instead, things like image preprocessing and model hyperparameters are key to training an accurate model.
@@
 % Both figures in one row
+% TODO: Move these to be under EDA in the right spots
 \begin{figure*}[!ht]
@@
 This is because the images without tumors have much less variance.
-It should be much easier for the model to detect tumor vs no tumor, so we do not expect the lower number of images in the no tumor class to have a negative effect on the model's ability to classify images as such.
+We expect it to be much easier for the model to detect "no tumor", and we do not think the lower number of images in the "no tumor" class will have a negative effect on the model's ability to classify images as such.
 However, if we do see poor results in that classification specifically, we can try changing all classes to train on 395 images, to see if that helps with the class-specific overfitting.
+Additionally we could also focus our data augmentation on the "no tumor" class, bringing its number of images up to the others.
 
@@
 It could be useful to test against more pituitary imagery to ensure the model can handle all variations of scans.
-We could perform data augmentation and create more data by rotating, zooming, and flipping our existing data.
+We could again perform data augmentation and create more pituitary data by rotating, zooming, and flipping our existing images.
 Testing against more data could give us more confidence in our model, if scores remain high.
@@
 
-We can see in Figure 2 that our dataset is comprised of images of varying sizes.
+We can see in Figure 2 that our dataset is comprised of images of many varying sizes.
 There seems to be a concentration around the 200x200 and 500x500 sizes.
 With this information, we know we will have to perform some image pre-processing to convert all images into the same size so the model has consistent input.
-A common image size is 256x256.
+We will provide our CNN with the conventional 224x244 image dimensions.
 
@@
 As mentioned in our dataset description, we don't have labels for our input features like a numerical dataset would have.
-We only have our images, which are technically a bunch of pixel intensity values, but it means we can't select certain features to train our dataset on.
+We only have our images, which are technically each a set of pixel intensity values, but it means we can't select certain features to train our dataset on.
 Instead, we can perform some image preprocessing and hyperparameter tuning to increase our model's accuracy.
 
-If we look at Figure 3 again, we see all of our images have different dimensions.
-As a preprocessing step, we will resize them all to be 256x256 pixels, so that our model has consistent input.
-We then normalize each pixel's value from its 3-channel RGB form.
-The resizing step is not as simple as it sounds though, we must be careful because shrinking or enlarging our images can create noise in our data.
-Based on future results, we might try a few different image sizes to see which is best.
+Looking at Figure 3 again, we see all of our images have different dimensions.
+As a preprocessing step, we will resize them all to be 224x224 pixels, so that our model has consistent input.
+The resizing step is not as simple as it sounds though, we must be careful because shrinking or enlarging our images can create unwanted noise in our data.
+After imagine resizing, we then normalize each pixel's value from its 3-channel RGB form.
 
 Our training dataset consists of 2870 MRI scans.
-If we want to give our model more images to train on, we can perform data augmentation.
+If we want to give our model more images to train on, we can perform data augmentation like mentioned previously.
 We can rotate, flip, and zoom our images to create 'artificial' data that will hopefully prevent overfitting and allow for our model to capture underlying patterns better.
@@
 
-We have already built an early prototype of a CNN model with 4 convolutional blocks, followed by fully connected layers using Keras Sequential API.
-We trained our model over 30 epochs, with each epoch giving us higher accuracy, peaking at 92.56\% after the last epoch.
-However, when tested against our testing data, the model performs quite poorly.
-We see an average accuracy of 42.13\%, indicating that the model is overfitting quite heavily on the training data.
+Our first early prototype was a CNN model with 4 convolutional blocks, followed by fully connected layers using the Keras Sequential API.
+We trained the model over 30 epochs, with each epoch giving us higher accuracy on the training data, peaking at 93\% after the last epoch.
+However, when performing predictions on our testing data, the model saw a big dip in performance.
+With an average accuracy of 42\%, our model was heavily overfitting on the training data.
 
-To improve upon our first model, we'll apply k-fold cross validation and tune hyperparameters such as dropout rate, number of layers/neurons, etc.
-We could start with some manual tuning, and eventually try implementing grid search to tune our hyperparameters.
-We also plan to use transfer learning to reduce the number of learned parameters and implicitly filter irrelevant features, which can greatly reduce overfitting risks.
-Going forward, we'll use PyTorch, which gives more flexibility in controlling forward pass, backpropagation, and optimization, to develop a better CNN model.
+% @Nam is this true? k-fold CV etc.?
+To improve upon our first model, we applied k-fold cross validation and tuned hyperparameters such as dropout rate, number of layers/neurons, etc.
+We implemented grid search to tune our hyperparameters.
+We also planned to use transfer learning to reduce the number of learned parameters and implicitly filter irrelevant features, which can greatly reduce overfitting risks.
+Going forward, we'd use PyTorch, which gave us more flexibility in controlling forward pass, backpropagation, and optimization, allowing us to develop a better CNN model.
 

### 2025-06-02 latex/draft.tex
more report drafting

@@
 \begin{abstract}
-This report documents the process of creating a machine learning model that classifies brain tumors using MRI brain scan imagery.
+This report documents the process of creating a machine learning model that classifies brain tumors given MRI brain scan imagery.
 This project saw the creation of three successful models using the following algorithms: Convolutional Neural Network, Support Vector Machine, and Random Forest.
@@
 \section{\large Introduction (Front page)}
-As a team of five young computer science students with no prior machine learning experience, we ambitously decided to tackle the real world challenge of tumor classification for our \textit{ECS 171 - Machine Learning} final project at UC Davis.
+As a team of five young computer science students with no prior machine learning experience, we ambitously decided to tackle the real world challenge of brain tumor classification for our \textit{ECS 171 - Machine Learning} final project at UC Davis.
 Through hours and hours of research and trial and error, we learned a lot, and we have a lot to share.
@@
 \subsection{\large Motivation}
+
+Brain tumors make up for... reference 9% TODO
 
@@
 
-Khan et al.
-(2021) examine convolutional neural network models for brain tumor classification, reporting classification accuracies ranging from 80\% to 98\%.
+In this article from 2021, authors Khan et al.
+examine convolutional neural network models for brain tumor classification, reporting classification accuracies ranging from 80\% to 98\%.
 The authors highlight two primary challenges in training such models: (1) the scarcity of accurately labeled data, which increases training complexity and cost, and (2) inconsistency in image dimensions, given that CNNs require fixed-size input.
@@
 
-In this article, Farooq, Anwar, Awais, and Rehman demonstrate their implementation of a 4-way classifier to classify different types of Alzheimer's disease, which is similar to our project since we are classifying our data into 4 sets.
-The authors began with data preprocessing and they utilized a technique known as skull stripping, which is the process of removing non-brain tissues like the skull or fat from MRI images.
-Since our data set consists of MRI images, we can implement skull stripping as well to improve signal-to-noise ratio and reduce input complexity, so it could help handle overfitting.
+In this article, Farooq, Anwar, Awais, and Rehman demonstrate their implementation of a four-way Alzheimer's disease classifier.
+Considering their model also takes MRI imagery as input and has four class target labels, we can learn a lot from their experiments.
+The authors began their model development with data preprocessing and a technique known as skull stripping, which is the process of removing non-brain tissues like the skull or fat from MRI imagery.
+We could implement skull stripping in our model to improve the signal-to-noise ratio and reduce input complexity, which could help against overfitting.
+
 Additionally, the article mentions the utilization of data augmentation to increase the data set; they simply flipped the images along the horizontal axis, which is valid due to the left and right symmetry of the brain.
@@
 The dataset used for this project comes from a group of brilliant students at Savitribai Phule Pune University in India.
-It can be found for free on Kaggle\textsuperscript{4}.
+It is an open source dataset that can be found for free on Kaggle\textsuperscript{4}.
 It is a well respected and reviewed dataset.
-The dataset consists of 3264 files, which are all images of MRI scans.
+The dataset consists of 3264 labeled MRI scan images.
 The images are divided into four categories.
@@
 Because each image in our dataset is already classified, our model's training is considered supervised, as opposed to unsupervised, where each training image would not have an associated classification.
+An unsupervised model would require a much different approach, most likely using different algorithms.
 
@@
     \includegraphics[width=0.49\textwidth]{ImagesPerClassTesting.png}
-    \caption{\large Number of images per class in the training (left) and testing (right) datasets}
+    \caption{\large The dataset has a consistent number of training images across classes, except for "no tumor".
+This should not be an issue because "no tumor" imagery has much less variance.
+For testing, the distribution of images is more varied, but this does not affect model performance, since the model is not trained on this data at all.}
     \label{fig:images-per-class}
@@
     \includegraphics[width=4in]{ImageDimensions.png}
-    \caption{\large Dimensions of the images in the dataset}
+    \caption{\large It is clear to see the images in the dataset are of many varying dimensions.
+We do see higher densities around the 200x200 and 500x500 dimensions, though.
+Regardless, such variation means image preprocessing and resizing is necessary before training.}
     \label{Dimensions of images in the dataset}
@@
 
+\bibitem{NBTS2022}
+Brain tumor facts. National Brain Tumor Society. (2024, February 20). https://braintumor.org/brain-tumors/about-brain-tumors/brain-tumor-facts/
+
 \end{thebibliography}

### 2025-06-02 latex/draft.tex
revised eval/conc

@@
 \section{\large Evaluation (1-2 pages)}
-In this project, we used three learning models (CNN, SVM, and RF) and compared the results of each. The dataset comprises four categories: glioma, meningioma, pituitary, and no tumor.
+In this project, three machine learning models—Convolutional Neural Networks (CNN), Support Vector Machines (SVM), and Random Forests (RF)—were trained and evaluated.
+In this section, the performance of these models are assessed using standard classification metrics, including accuracy, precision, recall, and F1 score.
+The dataset utilized for this evaluation consists of four classes: glioma, meningioma, pituitary tumor, and no tumor.
 
@@
 Diving into per-class specifics, we observe that for the label "glioma", CNN exhibited the highest precision (0.98) and recall (0.93), indicating strong identification capability with few false positives and a relatively low false negative rate.
-SVM matched CNN in precision (0.98) but had a lower recall (0.91), while RF also had a precision (0.98) but the lowest score in recall at 0.89.
+SVM matched CNN in precision (0.98) but had a lower recall (0.91), while RF also had a precision (0.98) but the lowest recall score at 0.89.
 This suggests that both SVM and RF are more likely to miss "glioma" cases compared to CNN while matching its precision.
@@
 
-In classifying "no tumor", RF achieved perfect precision (1.00), meaning that every time RF predicted "no tumor", the model was correct.
+In classifying "no tumor", RF achieved perfect precision (1.00), which means that every time RF predicted "no tumor", the model was correct.
 However, its recall (0.93) was the lowest among the three, pointing to a higher rate of false negatives.
@@
 
-For the label "pituitary", all three models performed strongly, with CNN leading in both precision and recall (0.99 each).
+For the label 'pituitary', all three models performed strongly, with CNN leading both in precision and recall (0.99 each).
 SVM and RF followed closely with both metrics at 0.98 and 0.97-0.98, respectively, indicating that all models are highly effective in classifying "pituitary" tumors.
@@
 \section{\large Conclusion (1/2 page)}
-The evaluation highlight CNN's strength in handling complex, high-dimensional image data, likely due to its hierarchical feature extraction capability.
-Its superior balance in both precision and recall across all tumor types supports its suitability for real-world clinical deployment where both false positives and false negatives carry significant risk.
+The above evaluation highlights CNN's strength in handling complex, high-dimensional image data, likely due to its hierarchical feature extraction capability.
+Its superior balance in both precision and recall in all tumor types supports its suitability for real-world clinical deployment, where false positives and false negatives carry significant risk.
 
-SVM performed comparably well, especially in the detection of meningioma and pituitary tumors, and may offer a lighter-weight alternative in environments with limited computational resources.
-However, it showed reduced precision in some classes, making it slightly less robust overall.
+SVM performed comparably well, especially in the detection of meningioma and pituitary tumors, and may offer a lighter alternative in environments with limited computational resources.
+However, it showed reduced precision in some classes, which made it slightly less robust overall.
 
 RF, while achieving remarkable precision, particularly for the "no tumor" class, exhibited reduced recall across multiple classes.
-This suggests a tendency to make fewer false positives but at the expense of missing actual tumor cases, which is a very critical and life-threatening issue in medical diagnostics.
+This suggests a tendency to make fewer false positives, at the expense of missing actual tumor cases, which is a very serious and life-threatening issue in medical diagnostics.
 
@@
 
-Richard: Richard has completed an article for literature review.
-Over the coming weeks, he will work towards exploratory data analysis (EDA) and preparing the dataset for analysis through data pre-processing.
+Richard: Richard contributed to the project by evaluating the performance of the three machine learning model using evaluation metrics such as accuracy, F1 score, precision, and recall.
+Based on these analyses, he provided the rationale for selecting the CNN model as the final model due to its superior performance relative to the alternatives.
+Additionally, he conducted a comprehensive literature review to summarize key developments in the field, which served as a foundational reference for the project.
+His contributions also included performing exploratory data analysis (EDA) and preparing the dataset for modeling through appropriate data preprocessing techniques.
 

### 2025-06-03 latex/draft.tex
Update draft.tex

@@
 
+Unlike accuracy, which can be misleading if one class dominates the dataset, the F1 score provides a balanced view of the model's ability to correctly identify each class without favoring either false positives or false negatives.
+This makes it especially useful in medical imaging applications, where both types of errors can have serious implications for patient diagnosis and treatment.
+
+The F1 score of 0.93 for glioma tumors.
+This score indicates that while the model performs well in identifying glioma cases, there is a notable trade-off between precision and recall, particularly with slightly reduced recall.
+This suggests the model may occasionally misclassify glioma tumors as other types or miss them altogether.
+
+The meningioma class has the lowest F1 score among all four classes with 0.92, indicating the greatest challenge for the model in this category.
+The score suggests a less favorable balance between precision and recall; either the model is producing more false positives, false negatives, or both.
+This could be due to visual overlap with other tumor types, particularly gliomas, or due to greater variability in meningioma imaging features.
+
+The F1 score of 0.97 for the no tumor class highlights the model’s strong performance in distinguishing healthy scans from non-healthy ones.
+The high precision and recall underlying this score indicate that the model rarely misclassifies healthy brains as tumors meaning there are a low amount of false positives and is also unlikely to miss the absence of a tumor or false negatives.
+This reliability is crucial in medical settings, where a false alarm can lead to unnecessary procedures, and a missed tumor can be life-threatening.
+
+With the highest F1 score of 0.98, the model shows excellent performance in detecting pituitary tumors.
+This suggests that the pituitary tumor class is the most distinct or consistently represented in the dataset, allowing the model to make confident predictions.
+High F1 scores for both precision and recall indicate that pituitary tumors are both correctly identified and rarely confused with other classes.
+This robustness may stem from more uniform imaging patterns or stronger contrast features in this tumor type.
+
 \section{\large Conclusion (1/2 page)}

### 2025-06-03 latex/draft.tex
report drafting on exp res

@@
 \section{\large Experimental Results (2 pages)}
-Talk about trying three different models, different pretrained models like ResNet, (the other one you tried first Nam I forgot what it was)
+Since deep learning is very efficient in detecting features from an image,
+and this characteristic is also beneficial to building other models using different base algorithms, 
+we have mainly focused on developing a Convolutional Neural Network (CNN) model with a high accuracy by iteratively refining both model architecture and training strategy.
+This section outlines the chronological development and iterative experimentation we have applied to finalize the set of models.
+While the detailed performance of each model will be elaborated on in the following sections, this discussion emphasizes the architecture refinement, technical adjustments, and discoveries made during the project that improved the overall performance over time.
+\subsection{\large Baseline CNN Model}
+Our very first model was built from scratch using TensorFlow’s Sequential API. 
+The architecture consisted of 4 convolution blocks with subsequent increases in the number of filters (kernels), 
+each of which is followed by a max pooling layer helping reduce the spatial dimensions. 
+The last layers of the model are characterized by a dense layer of 128 neurons with ReLU activation and an output layer of 4 neurons with Softmax activation for multi-class classification, each of which is accompanied by a dropout layer to mitigate overfitting.
+\
 
-Talk about fine-tuning, grid search for hyperparameter fine-tuning, etc.
+This model was trained over 20 epochs with categorical cross-entropy as the loss function and the Adam optimizer. 
+The model ended up with a training accuracy of 92.56\% and a validation accuracy of 76.67\%. 
+When tested on the testing set, the model only gave an accuracy of 42.13\%. The significant gap between training and testing performance implies signs of overfitting.
 
-Nam, Elaborate on this: "Alright guys, I finally figured out the problem.
-I've been using different pre-trained models but keep ending up with a lot of false negatives for the class glioma.
-I reviewed another notebook using the same dataset.
-Comparing it with our reference notebook, I notice that they both initially combined the training and testing folders, shuffled and split again.
-On the other hand, I'd been using the dataset as it is split on Kaggle.
-I believe there must be significant patterns in the testing folders on Kaggle that are not present in the training ones, especially the glioma-tumor class.
-So, merging them together and splitting again is the optimal solution here.
-Btw, I also decided to use Keras Function API instead of Sequential API (still same network architecture) and it turned out to be slightly better.
-I run the model several times and the accuracy is averagely around 92\% to 95\%"
+\subsection{\large Transfer Learning with ResNet50}
+In the next progression, we decided to employ the transfer learning technique, using the ResNet50 model,
+which is pre-trained on the ImageNet dataset. Instead of training the entire neural network from scratch, 
+we leveraged the architecture of ResNet50 as a feature extractor by eliminating its output layer and freezing all other layers. 
+The classification head was built on top of the pre-trained model, which served as a convolutional base, followed by a Global Average Pooling layer to compress spatial dimensions. 
+The classifier was completed by a fully connected dense layer of 128 neurons with ReLU activation and an output layer with softmax activation, which was also accompanied by a dropout rate of 0.3 to prevent overfitting.
+\
 
+The model was trained over 20 epochs. This transition yielded a significant improvement in validation accuracy (89.20\% compared to 76.67\% in the first model) 
+while maintaining a reasonable training accuracy (92.56\%).
+Motivated by this encouragement, we tried to experiment with selective fine-tuning the model by unfreezing the last 30 layers of the ResNet50 pre-trained model,
+allowing the weights of those layers to be adjusted during the backpropagation.
+With this technique, this model ended up with an accuracy of 97.53\%, 89.20\% and 79.19\% for the training, validation, and testing sets, respectively.
+At this stage, we also introduced new training practices such as learning rate management and early stopping, which shaped our training strategy for later experiments
 \section{\large Evaluation (1-2 pages)}

### 2025-06-03 latex/draft.tex
finished introduction

@@
 
-\title{Neural Networks for Neural Tumors}
+\title{Neural Networks for Neural Tumors\\
+\vspace{-0.6em}
+{\footnotesize Final Project for \textit{ECS 171 - Machine Learning} at UC Davis, with Professor Setareh Rafatirad}
+}
 
@@
   \end{tabular}
+  \vspace{-0.6em}
   \begin{tabular}{cc}
@@
 This project saw the creation of three successful models using the following algorithms: Convolutional Neural Network, Support Vector Machine, and Random Forest.
+Disclaimer: this project is for educational purposes only and not meant for medical use or research.
 \end{abstract}
@@
 \large 
-\section{\large Introduction (Front page)}
-As a team of five young computer science students with no prior machine learning experience, we ambitously decided to tackle the real world challenge of brain tumor classification for our \textit{ECS 171 - Machine Learning} final project at UC Davis.
+\section{\large Introduction (Complete)}
+As a team of five young computer science students with no prior machine learning experience, we ambitously decided to tackle the real world challenge of brain tumor classification.
 Through hours and hours of research and trial and error, we learned a lot, and we have a lot to share.
@@
 
-Brain tumors make up for... reference 9% TODO
+As of 2025, there are more than 120 different kinds of brain tumors\textsuperscript{10}.
+Population statistics gathered by braintumor.org from 2015 through 2019 led researchers to estimate that in 2023, over 94,000 people would be diagnosed with a primary brain tumor\textsuperscript{9}, in America alone.
+
+Luckily, 72\% of primary brain tumors are benign.
+While not completely harmless, benign tumors are much less fatal than malignant (cancerous) tumors.
+However, that still leaves a concerning 28\% of tumors that are malignant.
+Of those 28\%, an estimated 65\% of cases are proven to be fatal.
+Among the malignant tumors, glioblastoma is the most common, with a survival rate of only 6.9\%.
+These numbers are tragic, and at the moment, the only way we know to reduce them is through early detection.
 
@@
 
-Having a tool that could detect and identify tumors within seconds, and with high accuracy, would save a lot of time and money.
+A tool that could detect and identify brain tumors within seconds, and with high accuracy, would not only save a lot of time and money, but lives.
+Like mentioned above, early detection is absolutely key to increasing survival rates.
+Getting more people in and out of MRI machines and having their scans be automatically processed and analyzed would save countless lives.
 Not to mention the impact it would have on developing nations who might not have as many specially trained doctors that can readily perform MRI analysis.
-With lives on the line, early detection is absolutely key to diagnosing and treating these tumors.
+Luckily in the real world, hospitals are already adopting and developing these ideas.
 
@@
 
+\bibitem{Hopkins2025}
+Brain tumor types. Johns Hopkins Medicine. (2021, November 8). https://www.hopkinsmedicine.org/health/conditions-and-diseases/brain-tumor/brain-tumor-types
+
 \end{thebibliography}

### 2025-06-03 latex/draft.tex
Finish Experimental Results

@@
     \textit{UC Davis Computer Science} & \textit{UC Davis Computer Science} & \textit{UC Davis Computer Science} \\
-    jvogel@ucdavis.edu & ncnn@ucdavis.edu & jtnmo@ucdavis.edu \\
+    jvogel@ucdavis.edu & ncnng@ucdavis.edu & jtnmo@ucdavis.edu \\
     \vspace{0.6em}
@@
 At this stage, we also introduced new training practices such as learning rate management and early stopping, which shaped our training strategy for later experiments
+
+\subsection{\large Dataset Reorganization and ResNet101 upgrade}
+Motivated by the significant improvement in the previous stage, we had defined that transfer learning would be one of the most critical techniques in developing our CNN model.
+We decided to try different pre-trained models,
+from simpler models such as VGG16, VGG19 to more complicated ones in ResNet and EfficientNet families. 
+Even with either layer-wise fine-tuning or standard fine-tuning techniques applied to those models, 
+we had noticed the common trait, which is a persistent underperformance on the testing test (with an accuracy around 70.00\% to 80\%) 
+despite high accuracy in validation metrics (with an accuracy around 90\% to 93\%). This discrepancy prompted us to re-evaluate our dataset. 
+Initially, we had used the dataset split provided by Kaggle. We suspected that there might have been an imbalanced presentation across classes or feature distribution in the default split. 
+To resolve this issue, we decided to recombine the training and testing sets with shuffle, then split the entire dataset with a ratio consistent with that of the default split. \
+
+With available computational resources, we also upgraded to ResNet101, a larger architecture compared to ResNet50, 
+and transitioned to Keras Functional API to implement our model’s architecture for better flexibility and reusability. 
+We employed the same structure as in the ResNet50 setup.
+The model was initially trained over 20 epochs with all ResNet101 layers frozen, yielding a training accuracy of 91.72\% and validation accuracy of 85.07\%.
+When the fine-tuning technique was applied (with the last 40 layers unfrozen), the model reached 99.67\% training accuracy and 95.83\% validation accuracy after being trained over 4 additional epochs.
+The model also performed much better on the testing data, yielding 96.94\% accuracy with reasonably high f1-scores for each class label. \
+
+By employing ResNet101 with layer-wise fine-tuning and a better dataset split, 
+we ended up with a highly robust CNN model with consistent performance across training, validation, and testing data.
+
+\subsection{\large Non-deep Learning Models with feature extraction}
+To develop other classifiers that incorporate algorithms rather than deep learning, we decided to employ Support Vector Machine (SVM) and Random Forest Classifier (RF) models. 
+We leveraged the fine-tuned CNN model as the feature extraction for our selected dataset in order to obtain high accuracy while keeping the training process relatively simple. 
+More specifically, we extracted the output of the GlobalAveragePooling2D layer of the fine-tuned CNN model, which includes 2048-dimensional feature vectors, and that result would serve as the input for each non-deep learning model.
+In other words, our SVM and RF models were trained on a more semantically meaningful input space compared to raw pixels. \
+
+After conducting hyperparameter tuning, we obtained an overall accuracy of 95.66\% and 94.90\% for SVM and RF classifiers, respectively. 
+Although these two models were less efficient in distinguishing between glioma and meningioma compared to the CNN model, 
+other performance metrics were relatively high.
+In general, these results demonstrate that classical classifiers, when paired with feature extractors, can deliver highly competitive performance with significantly reduced training complexity
+
 \section{\large Evaluation (1-2 pages)}

### 2025-06-04 latex/draft.tex
more drafting, added model metric tables

@@
 \usepackage{algorithmic}
+\usepackage{booktabs}
 \usepackage{graphicx}
@@
     T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
-% \setlength{\parindent}{1em}
 % \setlength{\parskip}{0.8em}
@@
 }
-
-% old author block
-% \author{
-% \IEEEauthorblockN{\large Joe Vogel}
-% \IEEEauthorblockA{\textit{UC Davis Computer Science} \\
-% jvogel@ucdavis.edu}
-
-% \and
-% \IEEEauthorblockN{\large Nam Nguyen}
-% \IEEEauthorblockA{\textit{UC Davis Computer Science} \\
-% ncnn@ucdavis.edu}
-
-% \and
-% \IEEEauthorblockN{\large Jonathan Mo}
-% \IEEEauthorblockA{\textit{UC Davis Computer Science} \\
-% jtnmo@ucdavis.edu}
-
-% \and
-% \IEEEauthorblockN{\large Aryan Karnwal}
-% \IEEEauthorblockA{\textit{UC Davis Computer Science} \\
-% akarnwal@ucdavis.edu}
-
-% \and
-% \IEEEauthorblockN{\large Richard Ho}
-% \IEEEauthorblockA{\textit{UC Davis Computer Science} \\
-% rdho@ucdavis.edu}
-% }
 
@@
     jvogel@ucdavis.edu & ncnng@ucdavis.edu & jtnmo@ucdavis.edu \\
-    \vspace{0.6em}
+    \vspace{0.2em}
   \end{tabular}
@@
 
-As of 2025, there are more than 120 different kinds of brain tumors\textsuperscript{10}.
-Population statistics gathered by braintumor.org from 2015 through 2019 led researchers to estimate that in 2023, over 94,000 people would be diagnosed with a primary brain tumor\textsuperscript{9}, in America alone.
+As of 2025, there are more than 120 different kinds of brain tumors\textsuperscript{1}.
+Population statistics gathered by braintumor.org from 2015 through 2019 led researchers to estimate that in 2023, over 94,000 people would be diagnosed with a primary brain tumor\textsuperscript{2}, in America alone.
 
@@
 Among the malignant tumors, glioblastoma is the most common, with a survival rate of only 6.9\%.
-These numbers are tragic, and at the moment, the only way we know to reduce them is through early detection.
+These numbers are tragic, and at the moment, the best way we know to reduce them is through early detection.
 
@@
 Getting more people in and out of MRI machines and having their scans be automatically processed and analyzed would save countless lives.
-Not to mention the impact it would have on developing nations who might not have as many specially trained doctors that can readily perform MRI analysis.
+Not to mention the impact such a tool would have on developing nations who might not have as many specially trained doctors that can readily perform MRI analysis.
 Luckily in the real world, hospitals are already adopting and developing these ideas.
@@
 Using a convolutional neural network and an image dataset, we can train a model to accuractely classify brain tumors.
-With its complex neural structure, such a model can detect nuances and patterns in the scans that are otherwise hard to spot.
-This project used a dataset that focused on three brain tumor types: glioma, meningioma, pituitary.
-Along with the three tumor types, the model should also be able to classify scans as "no tumor".
-The aim for this model was to predict tumor classes with over 95\% accuracy.
+With its complex structure, a CNN can detect nuances and patterns in the scans that are otherwise hard to spot.
+
+This project focuses on three general brain tumor classes: glioma, meningioma, and pituitary.
+Each class contains many of the 120+ tumor types mentioned before.
+The model will also have a "no tumor" class.
+The aim for the model was to predict tumor classes with over 95\% accuracy.
 
@@
 
-\subsection{\large Brain Tumor Classification in MRI Image Using Convolutional Neural Network$^{1}$}
+\subsection{\large Brain Tumor Classification in MRI Image Using Convolutional Neural Network\textsuperscript{3}}
 
@@
 
-\subsection{\large A Deep CNN Based Multi-Class Classification of Alzheimer's Disease Using MRI$^{2}$}
+\subsection{\large A Deep CNN Based Multi-Class Classification of Alzheimer's Disease Using MRI\textsuperscript{4}}
 
@@
 
-\subsection{\large Brain Tumor Classification Using Convolutional Neural Networks$^{3}$}
+\subsection{\large Brain Tumor Classification Using Convolutional Neural Networks\textsuperscript{5}}
 
@@
 
-\section{\large Dataset Description (1 page)}
+\section{\large Dataset Description (Complete)}
 
 The dataset used for this project comes from a group of brilliant students at Savitribai Phule Pune University in India.
-It is an open source dataset that can be found for free on Kaggle\textsuperscript{4}.
-It is a well respected and reviewed dataset.
-The dataset consists of 3264 labeled MRI scan images.
-The images are divided into four categories.
-There are three different tumor types, and one set of “no tumor” scans.
-Because each image in our dataset is already classified, our model's training is considered supervised, as opposed to unsupervised, where each training image would not have an associated classification.
+It is an open source dataset that can be found for free on Kaggle\textsuperscript{6}.
+The dataset is reviewed and well respected.
+It consists of 3264 labeled MRI scan images.
+Like hinted at before, the images are divided into four categories: three different tumor types, and one set of “no tumor” scans.
+Because each image in our dataset is already labeled, our model's training is considered supervised, as opposed to unsupervised.
 An unsupervised model would require a much different approach, most likely using different algorithms.
@@
 Technically, each image is a set of numbers (pixel intensities), but the idea remains.
-We don't really have a standard feature selection process, or get to choose which features to train our model on.
+There is no standard feature selection process, or process to choose which features to train the model on.
 Instead, things like image preprocessing and model hyperparameters are key to training an accurate model.
 
-Preprocessing steps might include things like image resizing, cropping, normalization, and data augmentation (rotating, flipping, zooming, etc.).
-We can even consider techniques like 'skull stripping', used to isolate only the brain, since we do not care about anything else in the scan.
+As explained under 'Literature Review', preprocessing steps might include techniques like image resizing, cropping, normalization, data augmentation (rotating, flipping, zooming), and more.
+Going even further, the 'skull stripping' technique mentioned by Farooq et al.\textsuperscript{4} can be used to isolate the brain tissue in an MRI scan.
+This is helpful for training a model, since we do not care about the skull or fat in the scans, it is considered noise.
+We do not want the model to train on those "features" of the image.
 
-\section{\large EDA (1 page)}
+\section{\large Exploratory Data Analysis (1 page)}
+% TODO: Complete but need to fix image placement so that EDA is on one consistent page
 
@@
 This should not be an issue because "no tumor" imagery has much less variance.
-For testing, the distribution of images is more varied, but this does not affect model performance, since the model is not trained on this data at all.}
+For testing, the distribution of images is more varied, but this does not affect model performance, since the model is not trained on this data.}
     \label{fig:images-per-class}
@@
 
-As we can see in Figure 1 (left), there is an even distribution of images among each class (~825), except for the no tumor class (395).
-This is because the images without tumors have much less variance.
-We expect it to be much easier for the model to detect "no tumor", and we do not think the lower number of images in the "no tumor" class will have a negative effect on the model's ability to classify images as such.
-However, if we do see poor results in that classification specifically, we can try changing all classes to train on 395 images, to see if that helps with the class-specific overfitting.
-Additionally we could also focus our data augmentation on the "no tumor" class, bringing its number of images up to the others.
+As seen in the left histogram of \textit{Figure 1}, there is a relatively even distribution of data among each class in the training set, at around 825 images.
+However, the no tumor class has significantly less, with 395 images.
+We suspect this is because the scans without tumors have much less variance.
+And because of the reduced variance, it should be much easier for the model to detect "no tumor" scans.
+We did not think the lower number of images would have a negative effect on the model's ability to classify images as "no tumor", and we were right.
+However, if we did see poor results in that classification specifically, we could have tried training the model on 395 images for each class to see if that helped with the overfitting of the other classes.
+On the other hand, we could have performed data augmentation on the "no tumor" class, bringing its number of images up to the 825 range like the other classes.
+In hindsight, we did not have to apply either of these techniques.
 
-In our test dataset seen in Figure 1 (right), we have a much more uneven distribution among each class.
-This of course won't affect our model's ability, since it is not being trained on this data, but some results could be skewed considering the relatively low number of pituitary images.
-It could be useful to test against more pituitary imagery to ensure the model can handle all variations of scans.
-We could again perform data augmentation and create more pituitary data by rotating, zooming, and flipping our existing images.
-Testing against more data could give us more confidence in our model, if scores remain high.
+Shifting our focus toward the testing dataset, again looking at \textit{Figure 1}, but this time at the right histogram, we see that there is a much more uneven distribution among each class.
+This does not affect our model's performance, since it is not being trained on the testing data.
+However, we thought the results of the pituitary class could be skewed considering the relatively low number of pituitary, but it was also not a problem.
+If we had seen inconsitencies, it could have been useful to test against more pituitary imagery to ensure the model can handle all variations of scans.
+We could again have performed data augmentation to create more pituitary data.
+Testing against more data could give us more confidence in our model.
 
@@
 
-We can see in Figure 2 that our dataset is comprised of images of many varying sizes.
-There seems to be a concentration around the 200x200 and 500x500 sizes.
+Taking a look at Figure 2, we can see that the dataset is comprised of images of many varying sizes.
+There is a concentration around the 200x200 and 500x500 sizes.
 With this information, we know we will have to perform some image pre-processing to convert all images into the same size so the model has consistent input.
@@
 \section{\large Experimental Results (2 pages)}
-Since deep learning is very efficient in detecting features from an image,
-and this characteristic is also beneficial to building other models using different base algorithms, 
-we have mainly focused on developing a Convolutional Neural Network (CNN) model with a high accuracy by iteratively refining both model architecture and training strategy.
+
+Since deep learning is very efficient in detecting features from an image, and this characteristic is also beneficial to building other models using different base algorithms, we have mainly focused on developing a Convolutional Neural Network (CNN) model with a high accuracy by iteratively refining both model architecture and training strategy.
 This section outlines the chronological development and iterative experimentation we have applied to finalize the set of models.
 While the detailed performance of each model will be elaborated on in the following sections, this discussion emphasizes the architecture refinement, technical adjustments, and discoveries made during the project that improved the overall performance over time.
+
 \subsection{\large Baseline CNN Model}
-Our very first model was built from scratch using TensorFlow’s Sequential API. 
-The architecture consisted of 4 convolution blocks with subsequent increases in the number of filters (kernels), 
+
+Our very first model was built from scratch using TensorFlow’s Sequential API. The architecture consisted of 4 convolution blocks with subsequent increases in the number of filters (kernels), 
 each of which is followed by a max pooling layer helping reduce the spatial dimensions.
 The last layers of the model are characterized by a dense layer of 128 neurons with ReLU activation and an output layer of 4 neurons with Softmax activation for multi-class classification, each of which is accompanied by a dropout layer to mitigate overfitting.
-\
 
@@
 The model ended up with a training accuracy of 92.56\% and a validation accuracy of 76.67\%.
-When tested on the testing set, the model only gave an accuracy of 42.13\%. The significant gap between training and testing performance implies signs of overfitting.
+When tested on the testing set, the model only gave an accuracy of 42.13\%.
+The significant gap between training and testing performance implies signs of overfitting.
 
 \subsection{\large Transfer Learning with ResNet50}
-In the next progression, we decided to employ the transfer learning technique, using the ResNet50 model,
-which is pre-trained on the ImageNet dataset. Instead of training the entire neural network from scratch, 
-we leveraged the architecture of ResNet50 as a feature extractor by eliminating its output layer and freezing all other layers. 
+
+In the next progression, we decided to employ the transfer learning technique, using the ResNet50 model, which is pre-trained on the ImageNet dataset.
+Instead of training the entire neural network from scratch, we leveraged the architecture of ResNet50 as a feature extractor by eliminating its output layer and freezing all other layers.
 The classification head was built on top of the pre-trained model, which served as a convolutional base, followed by a Global Average Pooling layer to compress spatial dimensions.
@@
 
-The model was trained over 20 epochs. This transition yielded a significant improvement in validation accuracy (89.20\% compared to 76.67\% in the first model) 
-while maintaining a reasonable training accuracy (92.56\%).
-Motivated by this encouragement, we tried to experiment with selective fine-tuning the model by unfreezing the last 30 layers of the ResNet50 pre-trained model,
-allowing the weights of those layers to be adjusted during the backpropagation.
+The model was trained over 20 epochs.
+This transition yielded a significant improvement in validation accuracy (89.20\% compared to 76.67\% in the first model) while maintaining a reasonable training accuracy (92.56\%).
+Motivated by this encouragement, we tried to experiment with selective fine-tuning the model by unfreezing the last 30 layers of the ResNet50 pre-trained model, allowing the weights of those layers to be adjusted during the backpropagation.
 With this technique, this model ended up with an accuracy of 97.53\%, 89.20\% and 79.19\% for the training, validation, and testing sets, respectively.
@@
 
-\subsection{\large Dataset Reorganization and ResNet101 upgrade}
+\subsection{\large Dataset Reorganization and ResNet101 Upgrade}
+
 Motivated by the significant improvement in the previous stage, we had defined that transfer learning would be one of the most critical techniques in developing our CNN model.
-We decided to try different pre-trained models,
-from simpler models such as VGG16, VGG19 to more complicated ones in ResNet and EfficientNet families. 
-Even with either layer-wise fine-tuning or standard fine-tuning techniques applied to those models, 
-we had noticed the common trait, which is a persistent underperformance on the testing test (with an accuracy around 70.00\% to 80\%) 
-despite high accuracy in validation metrics (with an accuracy around 90\% to 93\%). This discrepancy prompted us to re-evaluate our dataset. 
-Initially, we had used the dataset split provided by Kaggle. We suspected that there might have been an imbalanced presentation across classes or feature distribution in the default split. 
-To resolve this issue, we decided to recombine the training and testing sets with shuffle, then split the entire dataset with a ratio consistent with that of the default split. \
+We decided to try different pre-trained models, from simpler models such as VGG16, VGG19 to more complicated ones in ResNet and EfficientNet families.
+Even with either layer-wise fine-tuning or standard fine-tuning techniques applied to those models, we had noticed the common trait, which is a persistent underperformance on the testing test (with an accuracy around 70.00\% to 80\%) despite high accuracy in validation metrics (with an accuracy around 90\% to 93\%).
+This discrepancy prompted us to re-evaluate our dataset.
+Initially, we had used the dataset split provided by Kaggle.
+We suspected that there might have been an imbalanced presentation across classes or feature distribution in the default split.
+To resolve this issue, we decided to recombine the training and testing sets with shuffle, then split the entire dataset with a ratio consistent with that of the default split.
+\
 
-With available computational resources, we also upgraded to ResNet101, a larger architecture compared to ResNet50, 
-and transitioned to Keras Functional API to implement our model’s architecture for better flexibility and reusability. 
+With available computational resources, we also upgraded to ResNet101, a larger architecture compared to ResNet50, and transitioned to Keras Functional API to implement our model's architecture for better flexibility and reusability.
 We employed the same structure as in the ResNet50 setup.
@@
 When the fine-tuning technique was applied (with the last 40 layers unfrozen), the model reached 99.67\% training accuracy and 95.83\% validation accuracy after being trained over 4 additional epochs.
-The model also performed much better on the testing data, yielding 96.94\% accuracy with reasonably high f1-scores for each class label. \
+The model also performed much better on the testing data, yielding 96.94\% accuracy with reasonably high f1-scores for each class label.
 
-By employing ResNet101 with layer-wise fine-tuning and a better dataset split, 
-we ended up with a highly robust CNN model with consistent performance across training, validation, and testing data.
+By employing ResNet101 with layer-wise fine-tuning and a better dataset split, we ended up with a highly robust CNN model with consistent performance across training, validation, and testing data.
 
 \subsection{\large Non-deep Learning Models with feature extraction}
+
 To develop other classifiers that incorporate algorithms rather than deep learning, we decided to employ Support Vector Machine (SVM) and Random Forest Classifier (RF) models.
@@
 More specifically, we extracted the output of the GlobalAveragePooling2D layer of the fine-tuned CNN model, which includes 2048-dimensional feature vectors, and that result would serve as the input for each non-deep learning model.
-In other words, our SVM and RF models were trained on a more semantically meaningful input space compared to raw pixels. \
+In other words, our SVM and RF models were trained on a more semantically meaningful input space compared to raw pixels.
 
 After conducting hyperparameter tuning, we obtained an overall accuracy of 95.66\% and 94.90\% for SVM and RF classifiers, respectively.
-Although these two models were less efficient in distinguishing between glioma and meningioma compared to the CNN model, 
-other performance metrics were relatively high.
+Although these two models were less efficient in distinguishing between glioma and meningioma compared to the CNN model, other performance metrics were relatively high.
 In general, these results demonstrate that classical classifiers, when paired with feature extractors, can deliver highly competitive performance with significantly reduced training complexity
@@
 \section{\large Evaluation (1-2 pages)}
+
+\begin{table}[h!]
+\centering
+\caption{CNN Metrics}
+\begin{tabular}{lcccc}
+\toprule
+\textbf{Class} & \textbf{Precision} & \textbf{Recall} & \textbf{F1-score} & \textbf{Support} \\
+\midrule
+glioma     & 0.98 & 0.93 & 0.95 & 109 \\
+meningioma & 0.93 & 0.97 & 0.95 & 105 \\
+no tumor   & 0.97 & 0.98 & 0.98 & 60  \\
+pituitary  & 0.99 & 0.99 & 0.99 & 118 \\
+\midrule
+\textbf{Accuracy}    &      &      & 0.97 & 392 \\
+\textbf{Macro avg}   & 0.97 & 0.97 & 0.97 & 392 \\
+\textbf{Weighted avg}& 0.97 & 0.97 & 0.97 & 392 \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{table}[h!]
+\centering
+\caption{SVM Metrics}
+\begin{tabular}{lcccc}
+\toprule
+\textbf{Class} & \textbf{Precision} & \textbf{Recall} & \textbf{F1-score} & \textbf{Support} \\
+\midrule
+glioma     & 0.98 & 0.91 & 0.94 & 109 \\
+meningioma & 0.89 & 0.97 & 0.93 & 105 \\
+no tumor   & 0.98 & 0.97 & 0.97 & 60  \\
+pituitary  & 0.98 & 0.98 & 0.98 & 118 \\
+\midrule
+\textbf{Accuracy}    &      &      & 0.96 & 392 \\
+\textbf{Macro avg}   & 0.96 & 0.96 & 0.96 & 392 \\
+\textbf{Weighted avg}& 0.96 & 0.96 & 0.96 & 392 \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{table}[h!]
+\centering
+\caption{RF Metrics}
+\begin{tabular}{lcccc}
+\toprule
+\textbf{Class} & \textbf{Precision} & \textbf{Recall} & \textbf{F1-score} & \textbf{Support} \\
+\midrule
+glioma     & 0.98 & 0.89 & 0.93 & 109 \\
+meningioma & 0.87 & 0.98 & 0.92 & 105 \\
+no tumor   & 1.00 & 0.93 & 0.97 & 60  \\
+pituitary  & 0.97 & 0.98 & 0.98 & 118 \\
+\midrule
+\textbf{Accuracy}    &      &      & 0.95 & 392 \\
+\textbf{Macro avg}   & 0.96 & 0.95 & 0.95 & 392 \\
+\textbf{Weighted avg}& 0.95 & 0.95 & 0.95 & 392 \\
+\bottomrule
+\end{tabular}
+\end{table}
+
 In this project, three machine learning models—Convolutional Neural Networks (CNN), Support Vector Machines (SVM), and Random Forests (RF)—were trained and evaluated.
@@
 \section{\large Conclusion (1/2 page)}
+
 The above evaluation highlights CNN's strength in handling complex, high-dimensional image data, likely due to its hierarchical feature extraction capability.
@@
 \section{\large Our Team (1 page)}
-The link to our GitHub\textsuperscript{8} repository for this project can be found in the references of this report.
 
-project roadmap? Reference the first one pager outline assignment we did on google doc
+The link to this project's GitHub\textsuperscript{10} repository can be found in the references of this report.
+The repository holds our source code, the dataset, this report, and some more supplementary files.
+Please feel free to fork the repository and mess around with it as you wish.
+
+Project roadmap? Reference the first one pager outline assignment we did on google doc
 %TODO --- roadmap?
+
 \subsection{\large Demonstration}
-include youtube video of our demo
+
+For those interested in seeing the model in action, a short five minute demonstration of our model can be found on YouTube\textsuperscript{11}.
+The demo briefly walks through how to use the model, how the frontend was developed, and how the model is invoked.
 
@@
 
-Joe: As the team leader, Joe has taken charge of organizing meetings, assigning tasks, and the general oversight over the project.
-In the coming weeks, Joe will be focusing on Exploratory Data Analysis, Literature Review, and Model Optimization.
+Joe: As the team leader, Joe took charge of logistics.
+He organized meetings, defined the roadmap for the project, and kept the team on track.
+He also lead the exploratory data anaylsis subproject, and was crucial to the creation of the final report.
 
@@
 \begin{thebibliography}{00}
+
+\bibitem{Hopkins2025}
+Brain tumor types. Johns Hopkins Medicine. (2021, November 8). https://www.hopkinsmedicine.org/health/conditions-and-diseases/brain-tumor/brain-tumor-types
+
+\bibitem{NBTS2022}
+Brain tumor facts. National Brain Tumor Society. (2024, February 20). https://braintumor.org/brain-tumors/about-brain-tumors/brain-tumor-facts/
 
@@
 
-\bibitem{NBTS2022}
-Brain tumor facts. National Brain Tumor Society. (2024, February 20). https://braintumor.org/brain-tumors/about-brain-tumors/brain-tumor-facts/
-
-\bibitem{Hopkins2025}
-Brain tumor types. Johns Hopkins Medicine. (2021, November 8). https://www.hopkinsmedicine.org/health/conditions-and-diseases/brain-tumor/brain-tumor-types
+\bibitem{Karnwal2025}
+A. Karnwal, "Brain Tumor Classification Application Demo". (2025), YouTube. https://www.youtube.com/watch?v=ECMGXgMrji0
 

### 2025-06-05 latex/draft.tex
cleaned up tables, removed weighted and macro averages, and support column
Update draft.tex

@@
 \caption{CNN Metrics}
-\begin{tabular}{lcccc}
+\begin{tabular}{lccc}
 \toprule
-\textbf{Class} & \textbf{Precision} & \textbf{Recall} & \textbf{F1-score} & \textbf{Support} \\
+\textbf{Class} & \textbf{Precision} & \textbf{Recall} & \textbf{F1-score} \\
 \midrule
-glioma     & 0.98 & 0.93 & 0.95 & 109 \\
-meningioma & 0.93 & 0.97 & 0.95 & 105 \\
-no tumor   & 0.97 & 0.98 & 0.98 & 60  \\
-pituitary  & 0.99 & 0.99 & 0.99 & 118 \\
+glioma     & 0.98 & 0.93 & 0.95 \\
+meningioma & 0.93 & 0.97 & 0.95 \\
+no tumor   & 0.97 & 0.98 & 0.98 \\
+pituitary  & 0.99 & 0.99 & 0.99 \\
 \midrule
-\textbf{Accuracy}    &      &      & 0.97 & 392 \\
-\textbf{Macro avg}   & 0.97 & 0.97 & 0.97 & 392 \\
-\textbf{Weighted avg}& 0.97 & 0.97 & 0.97 & 392 \\
+\textbf{Accuracy} \textbf{0.97} \\
 \bottomrule
@@
 \caption{SVM Metrics}
-\begin{tabular}{lcccc}
+\begin{tabular}{lccc}
 \toprule
-\textbf{Class} & \textbf{Precision} & \textbf{Recall} & \textbf{F1-score} & \textbf{Support} \\
+\textbf{Class} & \textbf{Precision} & \textbf{Recall} & \textbf{F1-score} \\
 \midrule
-glioma     & 0.98 & 0.91 & 0.94 & 109 \\
-meningioma & 0.89 & 0.97 & 0.93 & 105 \\
-no tumor   & 0.98 & 0.97 & 0.97 & 60  \\
-pituitary  & 0.98 & 0.98 & 0.98 & 118 \\
+glioma     & 0.98 & 0.91 & 0.94 \\
+meningioma & 0.89 & 0.97 & 0.93 \\
+no tumor   & 0.98 & 0.97 & 0.97 \\
+pituitary  & 0.98 & 0.98 & 0.98 \\
 \midrule
-\textbf{Accuracy}    &      &      & 0.96 & 392 \\
-\textbf{Macro avg}   & 0.96 & 0.96 & 0.96 & 392 \\
-\textbf{Weighted avg}& 0.96 & 0.96 & 0.96 & 392 \\
+\textbf{Accuracy} \textbf{0.96} \\
 \bottomrule
@@
 
-\begin{table}[h!]
+\begin{table}
 \centering
 \caption{RF Metrics}
-\begin{tabular}{lcccc}
+\begin{tabular}{lccc}
 \toprule
-\textbf{Class} & \textbf{Precision} & \textbf{Recall} & \textbf{F1-score} & \textbf{Support} \\
+\textbf{Class} & \textbf{Precision} & \textbf{Recall} & \textbf{F1-score} \\
 \midrule
-glioma     & 0.98 & 0.89 & 0.93 & 109 \\
-meningioma & 0.87 & 0.98 & 0.92 & 105 \\
-no tumor   & 1.00 & 0.93 & 0.97 & 60  \\
-pituitary  & 0.97 & 0.98 & 0.98 & 118 \\
+glioma     & 0.98 & 0.89 & 0.93 \\
+meningioma & 0.87 & 0.98 & 0.92 \\
+no tumor   & 1.00 & 0.93 & 0.97 \\
+pituitary  & 0.97 & 0.98 & 0.98 \\
 \midrule
-\textbf{Accuracy}    &      &      & 0.95 & 392 \\
-\textbf{Macro avg}   & 0.96 & 0.95 & 0.95 & 392 \\
-\textbf{Weighted avg}& 0.95 & 0.95 & 0.95 & 392 \\
+\textbf{Accuracy} \textbf{0.95} \\
 \bottomrule
@@
 Jonathan: Jonathan took charge of literature review and researched techniques that can be implemented to the model.
-In the next few weeks, Jonathan will be focusing on writing the proposed methodology for the final report.
+Jonathan was also responsible for writing the evaluation section that analyzed the performances of the models by talking about the evaluation metrics such as F1 scores, precision, and recall across the different classes.
 

### 2025-06-05 latex/draft.tex
more drafting, finished revising methodology and experimental results

@@
 \section{\large Introduction (Complete)}
+
 As a team of five young computer science students with no prior machine learning experience, we ambitously decided to tackle the real world challenge of brain tumor classification.
@@
 \subsection{\large Approach}
+
 A classification machine learning model is a great approach to this problem.
@@
 For testing, the distribution of images is more varied, but this does not affect model performance, since the model is not trained on this data.}
-    \label{fig:images-per-class}
+    \label{images-per-class}
 \end{figure*}
@@
     \centering
-    \includegraphics[width=4in]{ImageDimensions.png}
+    \includegraphics[width=5in]{ImageDimensions.png}
     \caption{\large It is clear to see the images in the dataset are of many varying dimensions.
@@
 With this information, we know we will have to perform some image pre-processing to convert all images into the same size so the model has consistent input.
-We will provide our CNN with the conventional 224x244 image dimensions.
 
-\section{\large Methodology (1-2 pages)}
+\section{\large Methodology (Complete)}
 
 \subsection{\large Feature Selection}
-As mentioned in our dataset description, we don't have labels for our input features like a numerical dataset would have.
-We only have our images, which are technically each a set of pixel intensity values, but it means we can't select certain features to train our dataset on.
+
+As mentioned in the dataset description, we don't have labels for our input features like a numerical dataset would have.
+We only have our images, which are technically each a set of pixel intensity values, but it means we can't select certain features to train our model on.
 Instead, we can perform some image preprocessing and hyperparameter tuning to increase our model's accuracy.
 
-Looking at Figure 3 again, we see all of our images have different dimensions.
-As a preprocessing step, we will resize them all to be 224x224 pixels, so that our model has consistent input.
-The resizing step is not as simple as it sounds though, we must be careful because shrinking or enlarging our images can create unwanted noise in our data.
+Looking at Figure 2 again, we see all of our images have different dimensions.
+As a preprocessing step, we will resize all images to be 224x224 pixels, the conventional CNN input size.
+The resizing step is not as simple as it sounds though, we must be careful not to create unwanted noise in our data.
 After imagine resizing, we then normalize each pixel's value from its 3-channel RGB form.
 
-Our training dataset consists of 2870 MRI scans.
-If we want to give our model more images to train on, we can perform data augmentation like mentioned previously.
-We can rotate, flip, and zoom our images to create 'artificial' data that will hopefully prevent overfitting and allow for our model to capture underlying patterns better.
-We also hope this will allow for our model to predict against a larger variety of testing data (i.e.
+The training dataset consists of 2870 MRI scans.
+To increase the size of the training dataset, we can perform data augmentation like mentioned previously.
+We can rotate, flip, and zoom the images to create `artificial' data that would hopefully reduce any overfitting and allow for the model to better capture underlying patterns.
+We also would expect this technique to allow for the model to predict against a larger variety of testing data (i.e.
 rotated images).
 
-The hyperparameters of our model are crucial to achieving a good fit. In our first iteration of the model, some of the hyperparameters we chose were as follows:
-%TODO --- fix table
-- Image size: 256x256 pixels		- Batch size: 32		- Epochs: 30	
-- Loss:  Categorical Cross Entropy 	- Metrics: Accuracy		- Dropout Rate: 0.25 and 0.50
-We plan on fine tuning these parameters with techniques like grid search as we continue to optimize our model.
+Along with input features, the hyperparameters of a model are crucial to achieving a good fit.
+The key hyperparameters for the CNN model were: batch size for gradient descent, training epochs, loss function, dropout rate, and a few others.
+We used grid search to find optimal parameters for all three of our models, more on that later.
 
@@
 
-Our first early prototype was a CNN model with 4 convolutional blocks, followed by fully connected layers using the Keras Sequential API.
-We trained the model over 30 epochs, with each epoch giving us higher accuracy on the training data, peaking at 93\% after the last epoch.
+Our first prototype model was a CNN with four convolutional blocks, followed by fully connected layers using the Keras Sequential API.
+We trained the model over 20 epochs, with each epoch giving us higher accuracy on the training data, peaking at 93\%.
 However, when performing predictions on our testing data, the model saw a big dip in performance.
-With an average accuracy of 42\%, our model was heavily overfitting on the training data.
+With an average accuracy of 42\%, we knew immediately that our model was heavily overfitting the training data.
 
-% @Nam is this true? k-fold CV etc.?
-To improve upon our first model, we applied k-fold cross validation and tuned hyperparameters such as dropout rate, number of layers/neurons, etc.
-We implemented grid search to tune our hyperparameters.
-We also planned to use transfer learning to reduce the number of learned parameters and implicitly filter irrelevant features, which can greatly reduce overfitting risks.
-Going forward, we'd use PyTorch, which gave us more flexibility in controlling forward pass, backpropagation, and optimization, allowing us to develop a better CNN model.
+To improve upon the first model, we applied k-fold cross validation and tuned hyperparameters such as dropout rate, number of layers, number of neurons, etc.
+This was the first time we used grid search to tune hyperparameters.
+It was also at this stage we decided that going forward, we would use transfer learning to reduce the number of learned parameters and implicitly filter irrelevant `features'.
+Transfer learning, which is the process of leveraging a pre-trained model to improve your own, can greatly reduce overfitting risks.
+From this point on, we used PyTorch.
+Pytorch gave us more flexibility in controlling forward pass, backpropagation, and optimization, allowing us to develop a better-fitting CNN model.
 
-\section{\large Experimental Results (2 pages)}
+\section{\large Experimental Results (Complete)}
 
-Since deep learning is very efficient in detecting features from an image, and this characteristic is also beneficial to building other models using different base algorithms, we have mainly focused on developing a Convolutional Neural Network (CNN) model with a high accuracy by iteratively refining both model architecture and training strategy.
-This section outlines the chronological development and iterative experimentation we have applied to finalize the set of models.
-While the detailed performance of each model will be elaborated on in the following sections, this discussion emphasizes the architecture refinement, technical adjustments, and discoveries made during the project that improved the overall performance over time.
+Neural networks like CNNs are very good at detecting features and learning patterns in data.
+Because of this, we can leverage deep learning to preprocess data, and then use that data to train other models.
+So far, this report has focused on developing a CNN model with high accuracy by iteratively refining both model architecture and training strategy.
+While CNNs are the most common algorithm choice for image classification, they are not the only option.
+Image classification can be done by various other algorithms, including support vector machine (SVM), and random forest (RF).
+
+While the detailed performance of each model we created will be discussed in the following sections, this section outlines the chronological development and iterative experimentation that lead us to our final models.
 
@@
 
-Our very first model was built from scratch using TensorFlow’s Sequential API. The architecture consisted of 4 convolution blocks with subsequent increases in the number of filters (kernels), 
-each of which is followed by a max pooling layer helping reduce the spatial dimensions.
-The last layers of the model are characterized by a dense layer of 128 neurons with ReLU activation and an output layer of 4 neurons with Softmax activation for multi-class classification, each of which is accompanied by a dropout layer to mitigate overfitting.
+Like mentioned before, our very first model was built from scratch using TensorFlow's Sequential API.
+The model consisted of four convolution blocks with subsequent increases in the number of filters (kernels).
+Each block was followed by a max pooling layer helping reduce the spatial dimensionality.
+Additionally, each hidden layer of the model was accompanied by a dropout layer to mitigate overfitting.
+The last hidden layer of the model was a dense layer of 128 neurons using the ReLU activation function.
+The output layer was four neurons using Softmax as the activation function.
+Softmax, given its consistent output between 0 and 1, is the standard choice of activation function for the output layer in multi-class classification.
 
-This model was trained over 20 epochs with categorical cross-entropy as the loss function and the Adam optimizer. 
-The model ended up with a training accuracy of 92.56\% and a validation accuracy of 76.67\%.
-When tested on the testing set, the model only gave an accuracy of 42.13\%.
-The significant gap between training and testing performance implies signs of overfitting.
+This model was trained over 20 epochs using the Adam optimizer and categorical cross-entropy as the loss function.
+And as mentioned before, the model ended up with a training accuracy of 93\% and a testing accuracy of 42\%, implying significant overfitting.
 
@@
 
-In the next progression, we decided to employ the transfer learning technique, using the ResNet50 model, which is pre-trained on the ImageNet dataset.
-Instead of training the entire neural network from scratch, we leveraged the architecture of ResNet50 as a feature extractor by eliminating its output layer and freezing all other layers.
-The classification head was built on top of the pre-trained model, which served as a convolutional base, followed by a Global Average Pooling layer to compress spatial dimensions.
-The classifier was completed by a fully connected dense layer of 128 neurons with ReLU activation and an output layer with softmax activation, which was also accompanied by a dropout rate of 0.3 to prevent overfitting.
-\
+In our next step, we decided to employ transfer learning.
+We started with the ResNet50 model, a pre-trained model trained on the ImageNet dataset.
+We leveraged the architecture of ResNet50 as a feature extractor by eliminating its output layer and freezing all other layers.
+The classification head was built on top of the pre-trained model, which served as a convolutional base.
+The base was followed by a Global Average Pooling layer to compress spatial dimensions.
+The classifier was again completed by a fully connected dense layer of 128 neurons with the ReLU activation function and a Softmax output layer.
+The layer of 128 neurons was accompanied by a dropout rate of 0.3 to help prevent overfitting.
 
 The model was trained over 20 epochs.
-This transition yielded a significant improvement in validation accuracy (89.20\% compared to 76.67\% in the first model) while maintaining a reasonable training accuracy (92.56\%).
-Motivated by this encouragement, we tried to experiment with selective fine-tuning the model by unfreezing the last 30 layers of the ResNet50 pre-trained model, allowing the weights of those layers to be adjusted during the backpropagation.
-With this technique, this model ended up with an accuracy of 97.53\%, 89.20\% and 79.19\% for the training, validation, and testing sets, respectively.
-At this stage, we also introduced new training practices such as learning rate management and early stopping, which shaped our training strategy for later experiments
+The transition to transfer learning yielded a significant improvement in validation accuracy (89\% compared to 76\% in the first model) while maintaining a reasonable training accuracy of 93\%.
+Motivated by this encouragement, we experimented with selective fine-tuning the model by unfreezing the last 30 layers of the ResNet50 pre-trained model, allowing the weights of those layers to be adjusted during backpropagation.
+With this adjustment, the model ended up with a training accuracy of 97\%, and a testing accuracy of 79\%.
+Although the model was still slightly overfitting, we had made huge improvements.
+At this stage, we also introduced new training practices such as the learning rate hyperparameter and early stopping, which shaped our training strategy for later experiments.
 
@@
 Motivated by the significant improvement in the previous stage, we had defined that transfer learning would be one of the most critical techniques in developing our CNN model.
-We decided to try different pre-trained models, from simpler models such as VGG16, VGG19 to more complicated ones in ResNet and EfficientNet families.
-Even with either layer-wise fine-tuning or standard fine-tuning techniques applied to those models, we had noticed the common trait, which is a persistent underperformance on the testing test (with an accuracy around 70.00\% to 80\%) despite high accuracy in validation metrics (with an accuracy around 90\% to 93\%).
+We decided to try different pre-trained models, from simpler models such as VGG16 and VGG19, to more complex ones in the ResNet and EfficientNet families.
+Even with either layer-wise fine-tuning or standard fine-tuning techniques applied to those models, we had noticed a common trait, persistent underperformance on the testing test (with an accuracy around 70\% to 80\%).
 This discrepancy prompted us to re-evaluate our dataset.
@@
 To resolve this issue, we decided to recombine the training and testing sets with shuffle, then split the entire dataset with a ratio consistent with that of the default split.
-\
 
-With available computational resources, we also upgraded to ResNet101, a larger architecture compared to ResNet50, and transitioned to Keras Functional API to implement our model's architecture for better flexibility and reusability.
+With limited available computational resources, we also upgraded our pre-trained model to ResNet101, a more complex model compared to ResNet50.
 We employed the same structure as in the ResNet50 setup.
-The model was initially trained over 20 epochs with all ResNet101 layers frozen, yielding a training accuracy of 91.72\% and validation accuracy of 85.07\%.
-When the fine-tuning technique was applied (with the last 40 layers unfrozen), the model reached 99.67\% training accuracy and 95.83\% validation accuracy after being trained over 4 additional epochs.
-The model also performed much better on the testing data, yielding 96.94\% accuracy with reasonably high f1-scores for each class label.
+The model was initially trained over 20 epochs with all ResNet101 layers frozen, yielding a training accuracy of 92\% and validation accuracy of 85\%.
+When the fine-tuning technique was applied (with the last 40 layers unfrozen), the model reached 99\% training accuracy and 96\% validation accuracy after being trained over four additional epochs.
+The model also performed much better on the testing data, yielding 97\% accuracy with reasonably high f1-scores for each class label.
+This was our breakthrough.
 
@@
 
-To develop other classifiers that incorporate algorithms rather than deep learning, we decided to employ Support Vector Machine (SVM) and Random Forest Classifier (RF) models.
+To develop other classifiers that incorporate algorithms rather than deep learning, we decided to employ Support Vector Machine (SVM) and Random Forest (RF) models.
 We leveraged the fine-tuned CNN model as the feature extraction for our selected dataset in order to obtain high accuracy while keeping the training process relatively simple.
@@
 
-After conducting hyperparameter tuning, we obtained an overall accuracy of 95.66\% and 94.90\% for SVM and RF classifiers, respectively.
-Although these two models were less efficient in distinguishing between glioma and meningioma compared to the CNN model, other performance metrics were relatively high.
-In general, these results demonstrate that classical classifiers, when paired with feature extractors, can deliver highly competitive performance with significantly reduced training complexity
+After conducting hyperparameter tuning, we obtained an overall accuracy of 96\% and 95\% for SVM and RF classifiers respectively.
+In general, these results demonstrate that classical classifiers, when paired with feature extractors, can deliver highly competitive performance for image classification, with significantly reduced training complexity.
 

### 2025-06-06 latex/draft.tex
added support column back to metric tables
Add Nam's contributions

@@
 \caption{CNN Metrics}
-\begin{tabular}{lccc}
+\begin{tabular}{lcccc}
 \toprule
-\textbf{Class} & \textbf{Precision} & \textbf{Recall} & \textbf{F1-score} \\
+\textbf{Class} & \textbf{Precision} & \textbf{Recall} & \textbf{F1-score} & \textbf{Support} \\
 \midrule
-glioma     & 0.98 & 0.93 & 0.95 \\
-meningioma & 0.93 & 0.97 & 0.95 \\
-no tumor   & 0.97 & 0.98 & 0.98 \\
-pituitary  & 0.99 & 0.99 & 0.99 \\
+glioma     & 0.98 & 0.93 & 0.95 & 109 \\
+meningioma & 0.93 & 0.97 & 0.95 & 105 \\
+no tumor   & 0.97 & 0.98 & 0.98 & 60 \\
+pituitary  & 0.99 & 0.99 & 0.99 & 118 \\
 \midrule
@@
 \caption{SVM Metrics}
-\begin{tabular}{lccc}
+\begin{tabular}{lcccc}
 \toprule
-\textbf{Class} & \textbf{Precision} & \textbf{Recall} & \textbf{F1-score} \\
+\textbf{Class} & \textbf{Precision} & \textbf{Recall} & \textbf{F1-score} & \textbf{Support} \\
 \midrule
-glioma     & 0.98 & 0.91 & 0.94 \\
-meningioma & 0.89 & 0.97 & 0.93 \\
-no tumor   & 0.98 & 0.97 & 0.97 \\
-pituitary  & 0.98 & 0.98 & 0.98 \\
+glioma     & 0.98 & 0.91 & 0.94 & 109 \\
+meningioma & 0.89 & 0.97 & 0.93 & 105 \\
+no tumor   & 0.98 & 0.97 & 0.97 & 60 \\
+pituitary  & 0.98 & 0.98 & 0.98 & 118 \\
 \midrule
@@
 \caption{RF Metrics}
-\begin{tabular}{lccc}
+\begin{tabular}{lcccc}
 \toprule
-\textbf{Class} & \textbf{Precision} & \textbf{Recall} & \textbf{F1-score} \\
+\textbf{Class} & \textbf{Precision} & \textbf{Recall} & \textbf{F1-score} & \textbf{Support} \\
 \midrule
-glioma     & 0.98 & 0.89 & 0.93 \\
-meningioma & 0.87 & 0.98 & 0.92 \\
-no tumor   & 1.00 & 0.93 & 0.97 \\
-pituitary  & 0.97 & 0.98 & 0.98 \\
+glioma     & 0.98 & 0.89 & 0.93 & 109 \\
+meningioma & 0.87 & 0.98 & 0.92 & 105 \\
+no tumor   & 1.00 & 0.93 & 0.97 & 60 \\
+pituitary  & 0.97 & 0.98 & 0.98 & 118 \\
 \midrule
@@
 
-Nam: Nam took charge of making the first prototype of our model.
-Having this model as a starting and reference point is absolutely crucial in the development of subsequent models.
-In the next few weeks, Nam will be researching different frameworks and libraries to develop other CNN models to see which best fit our dataset.
+Nam: Throughout the progression of the project, Nam was primarily responsible for the development and implementation of machine learning models. 
+Nam took over the experimental process to optimize each model, iteratively researching multiple architectures and frameworks to successfully develop models with different algorithms that are able to perform brain tumor detection at a relatively high accuracy.
+In general, Nam’s contributions encompassed every stage of model development, from initial prototyping to completion of each model prior to final deployment, ensuring both technical robustness and adaptability to team-driven refinements.
 

### 2025-06-07 latex/draft.tex
added Aryan contributions

@@
 
-Aryan: Aryan completed a literature review and analyzed research papers to identify techniques we can use to address the models overfitting.
-In the next few weeks Aryan will continue working to improve the models accuracy while also working on documentation and reports.
+Aryan: Aryan participated in group meetings, worked on literature reviews, helped the group brainstorm ideas and find the dataset, built and deployed the fullstack application, and created the demo video.
 

### 2025-06-07 latex/draft.tex
Completed report

@@
 
+\usepackage{balance}
 \usepackage{cite}
+\usepackage{tabularx}
 \usepackage{amsmath,amssymb,amsfonts}
@@
 
-As a team of five young computer science students with no prior machine learning experience, we ambitously decided to tackle the real world challenge of brain tumor classification.
+As a team of five young computer science students with no prior machine learning experience, we ambitously decided to tackle the real-world challenge of brain tumor classification.
 Through hours and hours of research and trial and error, we learned a lot, and we have a lot to share.
@@
 Not to mention the impact such a tool would have on developing nations who might not have as many specially trained doctors that can readily perform MRI analysis.
-Luckily in the real world, hospitals are already adopting and developing these ideas.
+Luckily in the real-world, hospitals are already adopting and developing these ideas.
 
@@
 Each class contains many of the 120+ tumor types mentioned before.
-The model will also have a "no tumor" class.
+The model will also have a ``no tumor" class.
 The aim for the model was to predict tumor classes with over 95\% accuracy.
@@
 Due to the non-numerical nature of image datasets, our input features come in a different form than they normally would with a numerical dataset.
-Instead of having numerically represented features such as 'age', 'blood pressure', and so on, our dataset only has images.
+Instead of having numerically represented features such as `age', `blood pressure', and so on, our dataset only has images.
 Technically, each image is a set of numbers (pixel intensities), but the idea remains.
@@
 
-As explained under 'Literature Review', preprocessing steps might include techniques like image resizing, cropping, normalization, data augmentation (rotating, flipping, zooming), and more.
-Going even further, the 'skull stripping' technique mentioned by Farooq et al.\textsuperscript{4} can be used to isolate the brain tissue in an MRI scan.
+As explained under `Literature Review', preprocessing steps might include techniques like image resizing, cropping, normalization, data augmentation (rotating, flipping, zooming), and more.
+Going even further, the `skull stripping' technique mentioned by Farooq et al.\textsuperscript{4} can be used to isolate the brain tissue in an MRI scan.
 This is helpful for training a model, since we do not care about the skull or fat in the scans, it is considered noise.
-We do not want the model to train on those "features" of the image.
+We do not want the model to train on those ``features" of the image.
 
@@
     \includegraphics[width=0.49\textwidth]{ImagesPerClassTesting.png}
-    \caption{\large The dataset has a consistent number of training images across classes, except for "no tumor".
-This should not be an issue because "no tumor" imagery has much less variance.
+    \caption{\large The dataset has a consistent number of training images across classes, except for ``no tumor".
+This should not be an issue because ``no tumor" imagery has much less variance.
 For testing, the distribution of images is more varied, but this does not affect model performance, since the model is not trained on this data.}
@@
 
-As seen in the left histogram of \textit{Figure 1}, there is a relatively even distribution of data among each class in the training set, at around 825 images.
+As seen in the left histogram of \textit{Figure 1} on the next page, there is a relatively even distribution of data among each class in the training set, at around 825 images.
 However, the no tumor class has significantly less, with 395 images.
 We suspect this is because the scans without tumors have much less variance.
-And because of the reduced variance, it should be much easier for the model to detect "no tumor" scans.
-We did not think the lower number of images would have a negative effect on the model's ability to classify images as "no tumor", and we were right.
+And because of the reduced variance, it should be much easier for the model to detect ``no tumor" scans.
+We did not think the lower number of images would have a negative effect on the model's ability to classify images as ``no tumor", and we were right.
 However, if we did see poor results in that classification specifically, we could have tried training the model on 395 images for each class to see if that helped with the overfitting of the other classes.
-On the other hand, we could have performed data augmentation on the "no tumor" class, bringing its number of images up to the 825 range like the other classes.
+On the other hand, we could have performed data augmentation on the ``no tumor" class, bringing its number of images up to the 825 range like the other classes.
 In hindsight, we did not have to apply either of these techniques.
@@
 
-Taking a look at Figure 2, we can see that the dataset is comprised of images of many varying sizes.
+Taking a look at Figure 2, again on the next page, we can see that the dataset is comprised of images of many varying sizes.
 There is a concentration around the 200x200 and 500x500 sizes.
@@
 When the fine-tuning technique was applied (with the last 40 layers unfrozen), the model reached 99\% training accuracy and 96\% validation accuracy after being trained over four additional epochs.
-The model also performed much better on the testing data, yielding 97\% accuracy with reasonably high f1-scores for each class label.
+The model also performed much better on the testing data, yielding 97\% accuracy with reasonably high F1-scores for each class label.
 This was our breakthrough.
@@
 
-\section{\large Evaluation (1-2 pages)}
+\section{\large Evaluation (Complete)}
 
-\begin{table}[h!]
+\begin{table*}[!ht]
 \centering
@@
 \end{tabular}
-\end{table}
+\end{table*}
 
-\begin{table}[h!]
+\begin{table*}[!ht]
 \centering
@@
 \end{tabular}
-\end{table}
+\end{table*}
 
-\begin{table}
+\begin{table*}[!ht]
 \centering
@@
 \end{tabular}
-\end{table}
+\end{table*}
 
-In this project, three machine learning models—Convolutional Neural Networks (CNN), Support Vector Machines (SVM), and Random Forests (RF)—were trained and evaluated.
-In this section, the performance of these models are assessed using standard classification metrics, including accuracy, precision, recall, and F1 score.
-The dataset utilized for this evaluation consists of four classes: glioma, meningioma, pituitary tumor, and no tumor.
+As previously mentioned, we trained three machine learning models for this project: a CNN, an SVM, and an RF model.
+In this section, the performances of these models are assessed using standard classification metrics such as accuracy, precision, recall, and F1-score.
 
-The CNN model demonstrated the highest overall accuracy at 96.68\%, followed closely by SVM at 95.66\%, and RF at 94.90\%.
-Macro and weighted F1-scores mirrored this trend, with CNN achieving 0.97, SVM 0.96, and RF 0.95 in both averages.
-These scores indicate that CNN not only achieves better overall classification performance but also maintains consistency across classes in terms of precision and recall.
+\vspace{1.5em}
 
-Diving into per-class specifics, we observe that for the label "glioma", CNN exhibited the highest precision (0.98) and recall (0.93), indicating strong identification capability with few false positives and a relatively low false negative rate.
-SVM matched CNN in precision (0.98) but had a lower recall (0.91), while RF also had a precision (0.98) but the lowest recall score at 0.89.
-This suggests that both SVM and RF are more likely to miss "glioma" cases compared to CNN while matching its precision.
+We can see from Tables 1, 2, and 3 that the CNN demonstrated the highest overall accuracy at 97\%, followed closely by the SVM at 96\%, and the RF at 95\%.
+The F1-scores of each model mirrored this trend, with the CNN achieving an average F1-score of 0.97, the SVM 0.96, and the RF 0.95.
+These scores indicate that the CNN not only achieves better overall classification performance, but also maintains consistency across classes in terms of precision and recall.
 
-In the case of "meningioma", while CNN had the best precision (0.93), SVM and RF had equal or superior recall (0.97 and 0.98 respectively) compared to CNN (0.97).
-SVM and RF had reduced precision (0.89 and 0.87) which is a notable difference from CNN's aforementioned score.
-This pattern indicates that SVM and RF are more sensitive in detecting "meningioma" but may yield more false positives than CNN.
+Diving into per-class metrics, again looking at the three tables, we observe that for the label ``glioma", all three models had an extremely high precision of 0.98, indicating very few false positives.
+In terms of recall, the CNN exhibited a score of 0.93, the SVM had a lower recall at 0.91, and the RF had the lowest recall at 0.89.
+These values highlight that both the SVM and the RF have higher false negative rates compared to the CNN for the ``glioma" class, meaning they classify glioma samples as other than glioma.
 
-In classifying "no tumor", RF achieved perfect precision (1.00), which means that every time RF predicted "no tumor", the model was correct.
-However, its recall (0.93) was the lowest among the three, pointing to a higher rate of false negatives.
-CNN balanced both precision (0.97) and recall (0.98) well, outperforming both alternatives in overall reliability for this class.
+In the case of the ``meningioma" class, while the CNN had the best precision at 0.93, the SVM and the RF had equal or superior recalls at 0.97 and 0.98 respectively, with the CNN at 0.97.
+The SVM and the RF had reduced precision scores at 0.89 and 0.87, which is a notable difference from the CNN's aforementioned score.
+This pattern indicates that the SVM and the RF saw more false positives, but equal or less false negatives.
+So if the goal was to reduce the number of ``meningioma" tumors classified as something else, the RF would be favorable by that 1\% difference we see in the recall scores.
 
-For the label 'pituitary', all three models performed strongly, with CNN leading both in precision and recall (0.99 each).
-SVM and RF followed closely with both metrics at 0.98 and 0.97-0.98, respectively, indicating that all models are highly effective in classifying "pituitary" tumors.
+In classifying ``no tumor", the RF achieved perfect precision of 1.00, which means that every time RF predicted ``no tumor", the model was correct.
+However, its recall at 0.93 was the lowest among the three, pointing to a higher rate of false negatives.
+The CNN balanced both precision and recall well, at 0.97 and 0.98 respectively, outperforming both alternatives in overall reliability for this class.
 
-Unlike accuracy, which can be misleading if one class dominates the dataset, the F1 score provides a balanced view of the model's ability to correctly identify each class without favoring either false positives or false negatives.
+For the label ``pituitary'', all three models performed strongly, with the CNN leading both in precision and recall both at 0.99.
+The SVM and RF followed closely with both metrics at 0.98 and 0.97-0.98, respectively, indicating that all models are highly effective in classifying ``pituitary" tumors.
+
+Unlike accuracy, which can be misleading if one class dominates the dataset, the F1-score provides a balanced view of the model's ability to correctly identify each class without favoring either false positives or false negatives.
 This makes it especially useful in medical imaging applications, where both types of errors can have serious implications for patient diagnosis and treatment.
 
-The F1 score of 0.93 for glioma tumors.
-This score indicates that while the model performs well in identifying glioma cases, there is a notable trade-off between precision and recall, particularly with slightly reduced recall.
-This suggests the model may occasionally misclassify glioma tumors as other types or miss them altogether.
+Because the CNN is the best general model of the three, we will take these next few paragraphs to analyze its F1-scores and their meanings.
+The model's F1-score of 0.95 for glioma tumors indicates that while the model performs well in identifying most glioma cases, there is a notable trade-off between precision and recall.
+The reduced recall value suggests the model may occasionally misclassify glioma tumors as other types or miss them altogether.
 
-The meningioma class has the lowest F1 score among all four classes with 0.92, indicating the greatest challenge for the model in this category.
-The score suggests a less favorable balance between precision and recall; either the model is producing more false positives, false negatives, or both.
-This could be due to visual overlap with other tumor types, particularly gliomas, or due to greater variability in meningioma imaging features.
+The ``meningioma" class also has an F1-score of .95, indicating a very similar story, except this time in reverse.
+The model has a lower precision compared to recall, meaning the model tends to classify tumors as ``meningioma'' when in fact they are not.
+This could be due to visual overlap with other tumor types, particularly gliomas, or due to greater variability in meningioma imaging.
 
-The F1 score of 0.97 for the no tumor class highlights the model’s strong performance in distinguishing healthy scans from non-healthy ones.
-The high precision and recall underlying this score indicate that the model rarely misclassifies healthy brains as tumors meaning there are a low amount of false positives and is also unlikely to miss the absence of a tumor or false negatives.
+The F1-score of 0.98 for the ``no tumor" class highlights the model's strong performance in distinguishing healthy scans from non-healthy ones.
+The high precision and recall underlying this score indicate that the model rarely misclassifies healthy brains as having a tumor (false negatives) and is also unlikely to miss the presence of a tumor (false positives).
 This reliability is crucial in medical settings, where a false alarm can lead to unnecessary procedures, and a missed tumor can be life-threatening.
 
-With the highest F1 score of 0.98, the model shows excellent performance in detecting pituitary tumors.
-This suggests that the pituitary tumor class is the most distinct or consistently represented in the dataset, allowing the model to make confident predictions.
-High F1 scores for both precision and recall indicate that pituitary tumors are both correctly identified and rarely confused with other classes.
+With the highest F1-score of 0.99, the model shows excellent performance in detecting pituitary tumors.
+This suggests that the ``pituitary" class is the most distinct or consistently represented in the dataset, allowing the model to make confident predictions.
+High scores for both precision and recall indicate that pituitary tumors are both correctly identified and rarely confused with other classes.
 This robustness may stem from more uniform imaging patterns or stronger contrast features in this tumor type.
 
-\section{\large Conclusion (1/2 page)}
+\section{\large Conclusion (Complete)}
 
-The above evaluation highlights CNN's strength in handling complex, high-dimensional image data, likely due to its hierarchical feature extraction capability.
+The above evaluation highlights the convolutional neural network's strength in handling complex, high-dimensional image data, likely due to its hierarchical feature extraction capability.
 Its superior balance in both precision and recall in all tumor types supports its suitability for real-world clinical deployment, where false positives and false negatives carry significant risk.
 
-SVM performed comparably well, especially in the detection of meningioma and pituitary tumors, and may offer a lighter alternative in environments with limited computational resources.
+The support vector machine model performed comparably well, especially in the detection of meningioma and pituitary tumors, and may offer a lighter alternative in environments with limited computational resources.
 However, it showed reduced precision in some classes, which made it slightly less robust overall.
 
-RF, while achieving remarkable precision, particularly for the "no tumor" class, exhibited reduced recall across multiple classes.
-This suggests a tendency to make fewer false positives, at the expense of missing actual tumor cases, which is a very serious and life-threatening issue in medical diagnostics.
+The random forest model, while achieving remarkable precision, particularly for the ``no tumor" class, exhibited reduced recall across multiple classes.
+This suggests a tendency to make fewer false positives, at the expense of missing actual tumor cases, which, as said before, is a very serious and life-threatening issue in medical diagnostics.
+
+Simply because of the extreme high-stakes nature of brain tumor classification, none of these three models we presented should be used in real-world medical diagnosis or research.
+But if you had to choose one, the CNN is the clear winner.
+It was expected from the start given the reputation of CNNs as image classifiers, but we must admit the results of the SVM and RF models were much closer than we expected.
 
@@
 
-The link to this project's GitHub\textsuperscript{10} repository can be found in the references of this report.
+Thank you for reading our paper.
+We had a great time researching, engineering, and writing up this project.
+The link to our GitHub\textsuperscript{10} repository can be found in the references of this report.
 The repository holds our source code, the dataset, this report, and some more supplementary files.
 Please feel free to fork the repository and mess around with it as you wish.
-
-Project roadmap? Reference the first one pager outline assignment we did on google doc
-%TODO --- roadmap?
 
@@
 \subsection{\large Roles and Contributions}
+\balance
 
-Joe: As the team leader, Joe took charge of logistics.
+\textbf{Joe}: As the team leader, Joe took charge of logistics.
 He organized meetings, defined the roadmap for the project, and kept the team on track.
-He also lead the exploratory data anaylsis subproject, and was crucial to the creation of the final report.
+He also led the exploratory data anaylsis subproject, and was the main author of the final report.
 
-Nam: Throughout the progression of the project, Nam was primarily responsible for the development and implementation of machine learning models. 
-Nam took over the experimental process to optimize each model, iteratively researching multiple architectures and frameworks to successfully develop models with different algorithms that are able to perform brain tumor detection at a relatively high accuracy.
-In general, Nam’s contributions encompassed every stage of model development, from initial prototyping to completion of each model prior to final deployment, ensuring both technical robustness and adaptability to team-driven refinements.
+\textbf{Nam}: Throughout the project, Nam was primarily responsible for the development and implementation of the machine learning models.
+Nam took over the experimental process of optimizing each model, researching architectures and frameworks to successfully develop models with different algorithms.
+In general, Nam's contributions encompassed every stage of model development, from initial prototyping to completion.
 
-Jonathan: Jonathan took charge of literature review and researched techniques that can be implemented to the model.
-Jonathan was also responsible for writing the evaluation section that analyzed the performances of the models by talking about the evaluation metrics such as F1 scores, precision, and recall across the different classes.
+\textbf{Jonathan}: Through literature review and careful research, Jonathan laid the groundwork for the project, finding out which techniques might be useful to us.
+Jonathan was also responsible for writing the evaluation section of the final report, analyzing the performances of each model by talking about the evaluation metrics such as F1-scores, precision, and recall.
 
-Richard: Richard contributed to the project by evaluating the performance of the three machine learning model using evaluation metrics such as accuracy, F1 score, precision, and recall.
-Based on these analyses, he provided the rationale for selecting the CNN model as the final model due to its superior performance relative to the alternatives.
-Additionally, he conducted a comprehensive literature review to summarize key developments in the field, which served as a foundational reference for the project.
-His contributions also included performing exploratory data analysis (EDA) and preparing the dataset for modeling through appropriate data preprocessing techniques.
+\textbf{Richard}: Richard contributed to the project by assessing the performance of the three machine learning models.
+Based on these analyses, he provided the rationale for selecting the CNN as the final model.
+Additionally, he conducted comprehensive literature review to summarize key developments in the field, which  like Jonathan's findings, served as a foundational reference for the project.
 
-Aryan: Aryan participated in group meetings, worked on literature reviews, helped the group brainstorm ideas and find the dataset, built and deployed the fullstack application, and created the demo video.
+\textbf{Aryan}: Aryan participated in group meetings, worked on literature reviews, and helped the group brainstorm ideas and find the dataset.
+Most importantly, he built and deployed the fullstack application, and recorded the demonstration video to go along with it.
+
+\subsection{\large Project Timeline / Roadmap}
+
+- April 19: First team meeting, finding a dataset
+
+- April 25: One-pager project description
+
+- May 01: Complete EDA
+
+- May 02: Initial model (42\% accuracy)
+
+- May 06: Mid-Quarter Report
+
+- May 15: Model improvements (~80\% accuracy)
+
+- May 18: Dataset reorganization (95\% accuracy)
+
+- May 20: Began drafting final report
+
+- May 22: SVM and RF models implemented
+
+- May 24: Frontend developed
+
+- June 07: Final report complete
 
@@
 \bibitem{Vogel2025}
-J. Vogel, N. Nguyen, J. Mo, A. Karnwal, R. Ho, "Neural Networks for Neural Tumors". GitHub. https://github.com/Coco501/Neural-Networks-
+J. Vogel, N. Nguyen, J. Mo, A. Karnwal, R. Ho, ``Neural Networks for Neural Tumors". GitHub. https://github.com/Coco501/Neural-Networks-
 for-Neural-Tumors
@@
 \bibitem{Karnwal2025}
-A. Karnwal, "Brain Tumor Classification Application Demo". (2025), YouTube. https://www.youtube.com/watch?v=ECMGXgMrji0
+A. Karnwal, ``Brain Tumor Classification Application Demo". (2025), YouTube. https://www.youtube.com/watch?v=ECMGXgMrji0
 

### 2025-06-07 latex/draft.tex
remove comments

@@
 \large 
-\section{\large Introduction (Complete)}
+\section{\large Introduction}
 
@@
 
-\section{\large Literature Review (Complete)}
+\section{\large Literature Review}
 
@@
 
-\section{\large Dataset Description (Complete)}
+\section{\large Dataset Description}
 
@@
 
-\section{\large Exploratory Data Analysis (1 page)}
-% TODO: Complete but need to fix image placement so that EDA is on one consistent page
-
-% EXAMPLE image in a column
-% \begin{figure}[!ht]
-%     \centering
-%     \includegraphics[width=2in]{coconut.png}
-%     \caption{this is a coconut}
-%     \label{test image coconut}
-% \end{figure}
-
-% EXAMPLE image across two columns
-% \begin{figure*}[!ht]
-%     \centering
-%     \includegraphics[width=3in]{coconut.png}
-%     \caption{this is a coconut}
-%     \label{coconut}
-% \end{figure*}
-
-% These two won't stay in the same row
-% \begin{figure*}[!ht]
-%     \centering
-%     \includegraphics[width=1in]{ImagesPerClassTraining.png}
-%     \caption{\large Images in the training dataset}
-%     \label{Number of images per class in the training dataset}
-% \end{figure*}
-
-% \begin{figure*}[!ht]
-%     \centering
-%     \includegraphics[width=1in]{ImagesPerClassTesting.png}
-%     \caption{\large Images in the testing dataset}
-%     \label{Number of images per class in the testing dataset}
-% \end{figure*}
-
-% Both figures in one row
-% TODO: Move these to be under EDA in the right spots
+\section{\large Exploratory Data Analysis}
 \begin{figure*}[!ht]
@@
 
-\section{\large Methodology (Complete)}
+\section{\large Methodology}
 
@@
 
-\section{\large Experimental Results (Complete)}
+\section{\large Experimental Results}
 
@@
 
-\section{\large Evaluation (Complete)}
+\section{\large Evaluation}
 
@@
 
-\section{\large Conclusion (Complete)}
+\section{\large Conclusion}
 
@@
 
-\section{\large Our Team (1 page)}
+\section{\large Our Team}
 
@@
 Based on these analyses, he provided the rationale for selecting the CNN as the final model.
-Additionally, he conducted comprehensive literature review to summarize key developments in the field, which  like Jonathan's findings, served as a foundational reference for the project.
+Additionally, he conducted comprehensive literature review to summarize key developments in the field which served as a foundational reference for the project.
 

### 2025-06-07 latex/draft.tex
final small revisions

@@
     T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
-% \setlength{\parskip}{0.8em}
 \begin{document}
@@
 
-As a team of five young computer science students with no prior machine learning experience, we ambitously decided to tackle the real-world challenge of brain tumor classification.
+As a team of five young computer science students with no prior machine learning experience, we ambitiously decided to tackle the real-world challenge of brain tumor classification.
 Through hours and hours of research and trial and error, we learned a lot, and we have a lot to share.
@@
 Hundreds of thousands of MRI brain scans are performed every day.
-With the scans being so intricate, it can sometimes be hard for even a trained professional to determine whether or not a patient has a brain tumor, and even harder to tell which kind.
+With the scans being so complex, it can sometimes be hard for even a trained professional to determine whether or not a patient has a brain tumor, and even harder to tell which kind.
 
 A tool that could detect and identify brain tumors within seconds, and with high accuracy, would not only save a lot of time and money, but lives.
-Like mentioned above, early detection is absolutely key to increasing survival rates.
+As mentioned above, early detection is absolutely key to increasing survival rates.
 Getting more people in and out of MRI machines and having their scans be automatically processed and analyzed would save countless lives.
@@
 A classification machine learning model is a great approach to this problem.
-Using a convolutional neural network and an image dataset, we can train a model to accuractely classify brain tumors.
+Using a convolutional neural network and an image dataset, we can train a model to accurately classify brain tumors.
 With its complex structure, a CNN can detect nuances and patterns in the scans that are otherwise hard to spot.
@@
 Their dataset is partitioned into training, validation, and testing sets, serving model training, hyperparameter tuning, and final evaluation, respectively.
-Image pre-processing is performed using Canny Edge Detection, an open-source computer vision technique.
+Image preprocessing is performed using Canny Edge Detection, an open-source computer vision technique.
 To address the limited dataset size, the authors employ data augmentation methods such as flipping, rotation, and brightness adjustments to enhance variability.
@@
 The eight-layer architecture discussed in the paper provides a useful baseline.
-The pre-processing techniques mentioned are also great starting points for us, and let us know what we should look out for.
+The preprocessing techniques mentioned are also great starting points for us, and let us know what we should look out for.
 Furthermore, we intend to implement data augmentation techniques to improve generalization on our small dataset.
@@
 In this article, Farooq, Anwar, Awais, and Rehman demonstrate their implementation of a four-way Alzheimer's disease classifier.
-Considering their model also takes MRI imagery as input and has four class target labels, we can learn a lot from their experiments.
+Considering their model also takes MRI imagery as input and has four class target labels, their methodology provides valuable insights.
 The authors began their model development with data preprocessing and a technique known as skull stripping, which is the process of removing non-brain tissues like the skull or fat from MRI imagery.
@@
 
-Additionally, the article mentions the utilization of data augmentation to increase the data set; they simply flipped the images along the horizontal axis, which is valid due to the left and right symmetry of the brain.
-This method increased their dataset from 355 MRI volumes to 38024 images.
-We can implement the same method in our project because increasing the sample size can combat overfitting, making our model more accurate.
+Additionally, the article mentions the utilization of data augmentation to increase the size of the dataset: they simply flipped the images along the horizontal axis, which is valid due to the left and right symmetry of the brain.
+This method grew their dataset from 355 MRI scans to 38,024.
+We can implement the same method in our project to increase the sample size, and combat overfitting, making our model more accurate.
 
@@
 This is helpful for training a model, since we do not care about the skull or fat in the scans, it is considered noise.
-We do not want the model to train on those ``features" of the image.
+We do not want the model to train on those features of the image.
 
@@
 There is a concentration around the 200x200 and 500x500 sizes.
-With this information, we know we will have to perform some image pre-processing to convert all images into the same size so the model has consistent input.
+With this information, we know we will have to perform some image preprocessing to convert all images into the same size so the model has consistent input.
 
@@
 The training dataset consists of 2870 MRI scans.
-To increase the size of the training dataset, we can perform data augmentation like mentioned previously.
+To increase the size of the training dataset, we can perform data augmentation as mentioned previously.
 We can rotate, flip, and zoom the images to create `artificial' data that would hopefully reduce any overfitting and allow for the model to better capture underlying patterns.
@@
 The key hyperparameters for the CNN model were: batch size for gradient descent, training epochs, loss function, dropout rate, and a few others.
-We used grid search to find optimal parameters for all three of our models, more on that later.
+Techniques like grid search can be used for hyperparameter tuning, more on that later.
 
@@
 This was the first time we used grid search to tune hyperparameters.
-It was also at this stage we decided that going forward, we would use transfer learning to reduce the number of learned parameters and implicitly filter irrelevant `features'.
+It was also at this stage we decided that going forward, we would use transfer learning to reduce the number of learned parameters and implicitly filter irrelevant features.
 Transfer learning, which is the process of leveraging a pre-trained model to improve your own, can greatly reduce overfitting risks.
-From this point on, we used PyTorch.
-Pytorch gave us more flexibility in controlling forward pass, backpropagation, and optimization, allowing us to develop a better-fitting CNN model.
 
@@
 
-Like mentioned before, our very first model was built from scratch using TensorFlow's Sequential API.
+As mentioned before, our very first model was built from scratch using TensorFlow's Sequential API.
 The model consisted of four convolution blocks with subsequent increases in the number of filters (kernels).
@@
 This model was trained over 20 epochs using the Adam optimizer and categorical cross-entropy as the loss function.
-And as mentioned before, the model ended up with a training accuracy of 93\% and a testing accuracy of 42\%, implying significant overfitting.
+Again, as mentioned before, the model ended up with a training accuracy of 93\% and a testing accuracy of 42\%, implying significant overfitting.
 
@@
 We decided to try different pre-trained models, from simpler models such as VGG16 and VGG19, to more complex ones in the ResNet and EfficientNet families.
-Even with either layer-wise fine-tuning or standard fine-tuning techniques applied to those models, we had noticed a common trait, persistent underperformance on the testing test (with an accuracy around 70\% to 80\%).
+Even with either layer-wise fine-tuning or standard fine-tuning techniques applied to those models, we had noticed a common trait: persistent underperformance on the testing test (with an accuracy around 70\% to 80\%).
 This discrepancy prompted us to re-evaluate our dataset.
@@
 In this section, the performances of these models are assessed using standard classification metrics such as accuracy, precision, recall, and F1-score.
+Accuracy describes the rate of correct predictions.
+Precision is the measurement of true positives against false positives.
+Recall measures true positives against false negatives.
+And F1-score is the harmonic mean of precision and recall.
 
@@
 Simply because of the extreme high-stakes nature of brain tumor classification, none of these three models we presented should be used in real-world medical diagnosis or research.
-But if you had to choose one, the CNN is the clear winner.
-It was expected from the start given the reputation of CNNs as image classifiers, but we must admit the results of the SVM and RF models were much closer than we expected.
+However, if you had to choose one, the CNN is the clear winner.
+It was expected from the start given the reputation of CNNs as image classifiers, but it must be said that the results of the SVM and RF models were much better than we expected.
 
@@
 He organized meetings, defined the roadmap for the project, and kept the team on track.
-He also led the exploratory data anaylsis subproject, and was the main author of the final report.
+He also led the exploratory data analysis subproject, and was the main author of the final report.
 

### 2025-06-07 latex/draft.tex
Final revisions

@@
 As of 2025, there are more than 120 different kinds of brain tumors\textsuperscript{1}.
-Population statistics gathered by braintumor.org from 2015 through 2019 led researchers to estimate that in 2023, over 94,000 people would be diagnosed with a primary brain tumor\textsuperscript{2}, in America alone.
+Population statistics gathered by braintumor.org from 2015 through 2019 led researchers to estimate that in 2023, over 94,000 people would be diagnosed with a primary brain tumor, in America alone\textsuperscript{2}.
 
@@
 The model will also have a ``no tumor" class.
-The aim for the model was to predict tumor classes with over 95\% accuracy.
+The aim for the model was to make predictions with over 95\% accuracy.
 
@@
 
-Additionally, the article mentions the utilization of data augmentation to increase the size of the dataset: they simply flipped the images along the horizontal axis, which is valid due to the left and right symmetry of the brain.
+Additionally, this article also mentions the utilization of data augmentation to increase the size of the dataset: they simply flipped the images along the horizontal axis, which is valid due to the left and right symmetry of the brain.
 This method grew their dataset from 355 MRI scans to 38,024.
@@
 Instead of having numerically represented features such as `age', `blood pressure', and so on, our dataset only has images.
-Technically, each image is a set of numbers (pixel intensities), but the idea remains.
+Technically, each image is a set of numbers (pixel intensities), but the idea remains the same.
 There is no standard feature selection process, or process to choose which features to train the model on.
@@
 
-As explained under `Literature Review', preprocessing steps might include techniques like image resizing, cropping, normalization, data augmentation (rotating, flipping, zooming), and more.
+As explained under the \textit{Literature Review} section, preprocessing steps might include techniques like image resizing, cropping, normalization, data augmentation (rotating, flipping, zooming), and more.
 Going even further, the `skull stripping' technique mentioned by Farooq et al.\textsuperscript{4} can be used to isolate the brain tissue in an MRI scan.
@@
 This does not affect our model's performance, since it is not being trained on the testing data.
-However, we thought the results of the pituitary class could be skewed considering the relatively low number of pituitary, but it was also not a problem.
-If we had seen inconsitencies, it could have been useful to test against more pituitary imagery to ensure the model can handle all variations of scans.
+However, we thought the results of the pituitary class could be skewed considering the relatively low number of scans, but it was also not a problem.
+If we had seen inconsitencies, it could have been useful to test against more pituitary imagery to ensure the model can handle more variations of scans.
 We could again have performed data augmentation to create more pituitary data.
@@
 
-Taking a look at Figure 2, again on the next page, we can see that the dataset is comprised of images of many varying sizes.
+Taking a look at \textit{Figure 2}, again on the next page, we can see that the dataset is comprised of images of many varying sizes.
 There is a concentration around the 200x200 and 500x500 sizes.
@@
 We only have our images, which are technically each a set of pixel intensity values, but it means we can't select certain features to train our model on.
-Instead, we can perform some image preprocessing and hyperparameter tuning to increase our model's accuracy.
+However, we can still perform image preprocessing and hyperparameter tuning to increase our model's accuracy.
 
-Looking at Figure 2 again, we see all of our images have different dimensions.
+Looking at \textit{Figure 2} again, we see all of our images have different dimensions.
 As a preprocessing step, we will resize all images to be 224x224 pixels, the conventional CNN input size.
@@
 The training dataset consists of 2870 MRI scans.
-To increase the size of the training dataset, we can perform data augmentation as mentioned previously.
+If needed, we can perform data augmentation as mentioned previously.
 We can rotate, flip, and zoom the images to create `artificial' data that would hopefully reduce any overfitting and allow for the model to better capture underlying patterns.
